% Abstracts
% ------------------------------------------------------------------
% Include an abstract in the language that the thesis is written in and in Finnish

% Abstract in English
% ------------------------------------------------------------------
\thesisabstract{english}{
Significance of organisation's ability to learn has been recently recognised in both academic and business world, especially in the context of new business ventures and innovation. Organisations that are able to learn faster than rivals and are thus better at adapting to changes in business environments are claimed to gain better competitive advantage. Recently, creativity and innovation abilities are related to be essential for employee's performance at work. Unpredictable, complex and uncertain environments require ability from both organisation and its employees to adapt to changes in creative ways that foster innovation. Thus, need for non-predictive approaches that support learning and growth in organisational and individual level exists. 

In this thesis, experimentation-driven approach for development is presented as a method for learning and building competitive advantage in an organisation. As a result from this study, factors affecting experimentation behaviour in organisation are presented and framework for organisational support for experimenting is created. 

Case study method was used in the study where client organisation was instructed to apply experimentation-driven approach during a six-week experimentation challenge aiming for employees to create novel ideas to develop their work and rapidly experiment those ideas. To study the factors affecting experimentation behaviour, an interpretive approach together with thematic analysis was used. The data consisted on 14 semi-structured interviews. 

Organisational and business environment being a complex system, various factors affect on the behaviour of employees and willingness to conduct experiments. Ability for holistic thinking and sensitivity for interdependencies is required. Experimentation behaviour is likely to be fostered by assuring safe environment for experimenting, supporting leadership behaviour, allocating resources for experimenting and careful experimentation design.
This thesis was done as a part of the two-year MINDexpe research project, undertaken by the MIND research group of Aalto University and funded by Tekes.
}

% Abstract in Finnish
% ------------------------------------------------------------------
\thesisabstract{finnish}{
Organisaation oppimiskyvyn merkitys on viime aikoina tunnistettu sekä akateemisessa että liiketoimintamaailmassa erityisesti uusien liikeyritysten ja innovaatioiden aloilla. Organisaatiot, jotka oppivat kilpailijoitaan nopeammin ovat siten parempia mukautumaan yrityselämän muutoksiin ja saavuttavan parempaa kilpailukykyä. Viime aikoina luovuus ja innovaatio on liitetty olennaiseksi osaksi työntekijän suorituskykyä. Ennalta arvaamattomat, monimutkaiset ja epävarmat ympäristöt vaativat sekä organisaatiolta että työntekijöiltä kykyä mukautua muutoksiin luovilla tavoilla, jotka edistävät innovaatioita. Näin on syntynyt tarve organisaation kehittämisen lähestymistavoille, jotka eivät pyri ennustamaan tulevaa, vaan tukevat ensisijaisesti organisaation sekä yksilön oppimista ja kasvua. 

Tässä diplomityössä kokeilemalla kehittäminen esitellään keinona oppia sekä lisätä organisaation kilpailukykyä. Työn tuloksena esitellään tekijöitä, jotka vaikuttavat kokeilevaan käyttäytymiseen organisaatiossa ja luodaan viitekehys tekijöistä, joilla organisaatio voi tukea kokeilemista. 

Tutkimuksessa käytettiin case-tutkimusta, jossa asiakasorganisaatio tutustutettiin kokeilemalla kehittämiseen kuuden viikon kokeilukilpailun kautta. Kilpailun tarkoitus oli luoda uusia ideoita työn parantamiseksi ja kokeilla niitä nopeasti. Tulkitsevaa tutkimusta yhdessä temaattisen analyysin kanssa käytettiin tutkimaan tekijöitä, jotka vaikuttavat kokeilemalla kehittämiseen. Empiriinen materiaali koostui 14 puolistrukturoidusta haastattelusta. 

Organisaatio ja liiketoimintaympäristö on monimutkainen systeemi, jossa monet tekjiät vaikuttavat työntekijöiden käytökseen ja halukkuuteen tehdä kokeiluja. Holistista ajattelua ja herkkyyttä tekijöiden välisille kokonaisriippuvuksille tarvitaan. Kokeilevaa käyttäytymistä voidaan tukea varmistamalla turvallinen ympäristö kokeilemiselle, kokeilemista tukevalla johtamisella, varaamalla riittävästi resursseja kokeilulle sekä huolellisella kokeilujen suunnittelulla. 

Tämä diplomityö tehdiin osana Aalto-yliopiston MIND-tutkimusryhmän kaksivuotista MINDexpe-tutkimushanketta, jonka rahoittaja oli Tekes. 
}
