\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\setcounter{secnumdepth}{5}

\begin{document}

\section{Change and learning in an organization}

Current and future business environment requires continuous innovation from organisations, meaning deploying the collective knowledge, skills and creative efforts of their employees (Dess and Picken, 2000). 

"As budgets are squeezed tighter and margins
of profit grow smaller, ideas are a precious
commodity and employees who produce them
become sought-after resources. Within such a
competitive business environment
companies nowadays increasingly strive to
become innovative organisations." (Andriopoulos and Lowe, 2000)

"Change. The rapid technological
advancements and the fierce competition
for market share have contributed to the
unprecedented increasing pace of change.
Therefore, organisations should be ready
to rearrange their resources to meet the
new demands." (Andriopoulos and Lowe, 2000)

Wide access to the information has put tremendous pressure on today's business and companies to increase their efficiency and effectiveness. Concurrently, demand for creative endeavours has rose in order to improve and develop products and processes. (Andriopoulos and Lowe, 2000; Cummings and Oldham, 1996)

"In today?s innovation-driven economy, 
understanding how to generate great ideas 
is an urgent managerial priority. And that 
calls for major doses of creativity. But many 
leaders assume creativity is too elusive and 
intangible to be managed." (Amabile and Khaire 2008)

Although currently creativity and creative processes of an individual at work are rather well recognised and essential, even more focus should be put on organisations' ability to mobilise creative actions of employees to create novel, socially valued products or services and more efficient ways of working (Mumford and Gustafson, 1988). In other words, creative actions of an employee are not worthwhile to an organisation when not coordinated or harnessed to yield organisational-level outcomes (Jung et al. 2003) 

"Increasingly, creativity has become valued across a variety of tasks, occupations, and industries. In today's fast-paced dynamic work environment, managers continue to realize that to remain competitive they need their employees to be actively involved in their work and trying to generate novel and appropriate products, processes, and approaches. Although the level of creativity required and the importance of creativity can differ depending on the tasks or job in question, most managers would agree that there is room, in almost every job, for employees to be more creative. Further, because individual creativity provides the foundation for organizational creativity and innovation (Amabile, 1988), and these have been linked to firm performance and survival (Nystrom, 1990), it is important, if not critical, that employees are creative in their work." (Shalley and Gilson 2004) 


According to Schein (2010), concept of culture refers to and helps to explain some seemingly incomprehensible and irrational aspects of what is going on in groups, organisations and other kinds of social units, that share history. Schein divides culture into three levels: artefacts (visible and feelable structures and processes and observed behaviour), espoused beliefs and values (ideals, goals, aspirations, ideologies and rationalisations) and basic underlying assumptions (unconscious, taking-for-granted beliefs and values). Climate of the group should not be mixed with culture of the group, it should rather be considered among artefacts. However, essential point of view Schein (2010) provides is how culture in organisation or group level is easy to observe yet very difficult to decipher. Put in other words: researchers are able to observe and make remarks on what they see and feel, yet they are unable to reconstruct the deeper meaning of those observations to the group. Cultural analysis and understanding of dynamics of a group should begin in observing and asking members the norms, values and rules that shape practicalities of work in day-to-day level.


\subsection{Learning as competitive advantage and way to survive}

Understanding change analytically in the turbulent world appears challenging. Change being hectic and fast calls for different skills and strategy than before. Only when change is understood can it be managed, and in order to survive new perspective and understanding towards change is required from an organisation. 

In the changing environment tolerance for uncertainty is needed, and while future can not be predicted, forecasting is a usable method in order to cope with the anxiousness resulting from the uncertainty. Furthermore, even more emphasis should be put on ability to learn and adapt to changes. 

"In this study, my focus on learning behavior and its accom- 
panying risk made the interpersonal context especially sa- 
lient; however, the need for learning in work teams is likely 
to become increasingly critical as organizational change and 
complexity intensify. Fast-paced work environments require 
learning behavior to make sense of what is happening as 
well as to take action. With the promise of more uncertainty, 
more change, and less job security in future organizations, 
teams are in a position to provide an important source of 
psychological safety for individuals at work. The need to ask 
questions, seek help, and tolerate mistakes in the face of 
uncertainty-while team members and other colleagues 
watch-is probably more prevalent in companies today than 
in those in which earlier team studies were conducted" (Edmondson 1999)

\subsection{Change in organisational level}
Hammer and Champy (1993) have summarised aspects of change in organisational environment, beginning from the change in organisational structure; from functional departments to process teams. Work tasks change from simple and detailed tasks to multi-dimensional knowledge work while employees are becoming more autonomous instead of strict control. Furthermore, instead of educating, focus is in the learning of an employee, and evaluation of work will change from operations to outcomes. Knowledge and capability are preferred over single performance and values change to more productive behaviour than over-protective. Superiors turn from leaders of the work to coaches and hierarchical organisational structures turn lower while managers focus on leadership instead of task management. 

According to Arie de Geus (1997) the only one can maintain company's competitive advantage is to make sure the company is able to learn faster than rivals. Generally organisations are considered as machines, yet recently more emphasis has been put on organisations as living organisms. When considered as machine, organisational model is mechanic and simple, which purpose is to gain profit. Whereas, organisation as a living organism is a whole-systemic model, and organisations are considered as place which has deeper, permanent meaning offering people the opportunity to grow and fulfil themselves while earning money.

Liable vision of the future focuses on the latter perspective of organisations, where learning and renewal form the essence of being. 

According to Huy and Minztberg (2003), organisations learn best through small experiments and trying out new things, and the closer and more related experimentations are to customers and customer interfaces, the more can be learned. 

\subsection{Organizational learning}

"Organizational learning is presented in the literature in two 
different ways: some discuss learning as an outcome; others 
focus on a process they define as learning. For example, 
Levitt and March (1 988: 320) conceptualized organizational 
learning as the outcome of a process of organizations "en- 
coding inferences from history into routines that guide be- 
havior"; in contrast, Argyris and Sch6n (1978) defined learn- 
ing as a process of detecting and correcting error. In this 
paper I join the latter tradition in treating learning as a pro- 
cess and attempt to articulate the behaviors through which 
such outcomes as adaptation to change, greater understand- 
ing, or improved performance in teams can be achieved. For 
clarity, I use the term "learning behavior" to avoid confusion 
with the notion of learning outcomes. " (Edmondson 1999)

"The conceptualization of learning as a process has roots in 
the work of educational philosopher John Dewey, whose 
writing on inquiry and reflection (e.g., Dewey, 1956) has had 
considerable influence on subsequent learning theories (e.g., 
Kolb, 1984; Schbn, 1983). Dewey (1956) described learning 
as an iterative process of designing, carrying out, reflecting 
upon, and modifying actions, in contrast to what he saw as 
the human tendency to rely excessively on habitual or auto- 
matic behavior. Similarly, I conceptualize learning at the 
group level of analysis as an ongoing process of reflection 
and action, characterized by asking questions, seeking feed- 
back, experimenting, reflecting on results, and discussing 
errors or unexpected outcomes of actions. For a team to dis- 
cover gaps in its plans and make changes accordingly, team 
members must test assumptions and discuss differences of 
opinion openly rather than privately or outside the group. I 
refer to this set of activities as learning behavior, as it is 
through them that learning is enacted at the group level." (Edmonsdon 1999)

"This conceptualization is consistent with a definition of group 
learning proposed recently by Argote et al. (2001) as both processes and outcomes of group interaction 
activities through which individuals acquire, share, and com- 
bine knowledge, but it focuses on the processes and leaves 
outcomes of these processes to be investigated separately. " (Edmondson 1999)

"The management literature encompasses related discussions 
of learning, for example, learning as dependent on attention 
to feedback (Schon, 1983), experimentation (Henderson and 
Clark, 1990), and discussion of failure (Sitkin, 1992)" 

"Similarly, because errors provide a source of information about performance by revealing that something did not work as planned, the ability to discuss them productively has been associated with organizational effectiveness (Michael, 1976; Sitkin, 1992; Schein, 2010) (Edmondson)

"The construct has roots in early research on organiza- 
tional change, in which Schein and Bennis (1965) discussed 
the need to create psychological safety for individuals if they 
are to feel secure and capable of changing. Team psycho- 
logical safety is not the same as group cohesiveness, as re- 
search has shown that cohesiveness can reduce willingness 
to disagree and challenge others' views, such as in the phe- 
nomenon of groupthink (Janis, 1982), implying a lack of inter- 
personal risk taking. The term is meant to suggest neither a 
careless sense of permissiveness, nor an unrelentingly posi- 
tive affect but, rather, a sense of confidence that the team 
will not embarrass, reject, or punish someone for speaking 
up. This confidence stems from mutual respect and trust 
among team members. " (Edmondson 1999)

"The importance of trust in groups and organizations has long 
been noted by researchers (e.g., Golembiewski and Mc- 
Conkie, 1975; Kramer, 1999). Trust is defined as the expec- 
tation that others' future actions will be favorable to one's 
interests, such that one is willing to be vulnerable to those 
actions (Mayer, Davis, and Schoorman, 1995; Robinson, 
1996). Team psychological safety involves but goes beyond 
interpersonal trust; it describes a team climate characterized 
by interpersonal trust and mutual respect in which people 
are comfortable being themselves. "

From the data rose various factors affecting experimentation behaviour, and the literature review revealed how similar factors are related to learning in an organisation and organisational learning. 

Through the understanding of organisational learning real growth and support for learning can be offered. Learning process needs to be understood all from organisational, team and individual perspective. 

Organizational learning is approached from two different perspectives in literature. On the one hand, learning is considered as an outcome, and on the other it is considered as a process (Edmondson 1999). In the first perspective organizational learning is referred to be an outcome of a process of organisations "encoding interferences from history into routines that guide behaviour (Levitt and March 1988). Whereas process perspective define learning as a process of continuous trial and error (Argyris and Schön 1978). Further definitions of learning are presented in chapter zz. In this thesis, learning is considered as the latter tradition of learning, which allows the growth and improved performance of individuals and organisations.

Organizational learning differs from individual or team learning. Organizational learning occurs through the shared knowledge, insights and approaches of the employees of an organisation. Secondly, organisational learning is based on prior knowledge and experience, the memory of organisation, which consists of the ways of working, processes and instructions of an organisation. Even though individual and team learning are highly related to organisational learning, it is not the sum of the previously mentioned. (Sydänmaanlakka 2007)

Various definitions of learning organisations have been presented. One of the most famous definitions is from Peter Senge (1991), who describes learning organisation as follows: "Learning organisation is an organisation, where people are able to constantly develop and achieve intended results; where new ways of thinking are born and where people share goals and learn together."

"Learning, analysis, imitation, regeneration, and technological change are major
components of any effort to improve organizational performance and strengthen
competitive advantage. Each involves adaptation and a delicate trade-off between
exploration and exploitation. The present argument has been that these trade-offs are
affected by their contexts of distributed costs and benefits and ecological interaction.
The essence of exploitation is the refinement and extension of existing competences,
technologies, and paradigms. Its retums are positive, proximate, and predictable. The
essence of exploration is experimentation with new altematives. Its returns are
uncertain, distant, and often negative. Thus, the distance in time and space between
the locus of leaming and the locus for the realization of returns is generally greater in
the case of exploration than in the case of exploitation, as is the uncertainty." (March 1991)

"an organization's structure can play a critical role in enhancing or hindering creativity." (Shalley and Gilson 2004) 

According to Sternberg et al. (1997) a company can enhance its creative skills by focusing on six resources: knowledge, intellectual abilities, thinking styles, motivation, personality and environment. Sternberg et al. (1997) argues that too much information of may hinder change and be seen as rigidity in thinking. Thus, one should not over-weight the criticism of senior people in an organisation, and at least consider the chance for rigid thinking and intolerance for change. 
In addition, needs to be noted and understood that employees' thinking styles are shaped through what is rewarded meaning, that if organisational environment rewards well-behaving and instruction-following thinking style and action, employees tend to implement their style to that. We are urged to adapt to organisational style and fit in, and when this is not possible, people tend to leave. 

\subsection{Meaning of a team}

"A growing reliance on teams in changing and uncertain orga- 
nizational environments creates a managerial imperative to 
understand the factors that enable team learning. Although 
much has been written about teams and about learning in 
organizations, our understanding of learning in teams re- 
mains limited" (Edmondson 1999)

"--little research has been 
done to understand the factors that influence learning behav- 
ior in ongoing teams in real organizations" (Edmondson 1999)

"Studies of work teams in a variety of organizational settings 
have shown that team effectiveness is enabled by structural 
features such as a well-designed team task, appropriate 
team composition, and a context that ensures the availability 
of information, resources, and rewards (Hackman, 1987). "

"Many researchers have concluded that structure and design, 
including equipment, materials, physical environment, and 
pay systems, are the most important variables for improving 
work-team performance (Goodman, Devadas, and Hughson, 
1988; Campion, Medsker, and Higgs, 1993; Cohen and Led- 
ford, 1994) and have argued against focusing on interper- 
sonal factors (e.g., Goodman, RavIin, and Schminke, 1987). 
According to this research, organization and team structures 
explain most of the variance in team effectiveness. 
In contrast, organizational learning research has emphasized 
cognitive and interpersonal factors to explain effectiveness, 
showing, for example, that individuals' tacit beliefs about in- 
terpersonal interaction inhibit learning behavior and give rise 
to ineffectiveness in organizations (e.g., Argyris, 1993). This 
cognitive emphasis takes different forms." (Edmondson 1999)

"This paper takes a different approach 
to understanding learning in organizations by examining to 
what extent and under what conditions learning occurs natu- 
rally in organizational work groups."(Edmondson 1999)

"Organizational work teams are groups that 
exist within the context of a larger organization, have clearly 
defined membership, and share responsibility for a team 
product or service (Hackman, 1987; Alderfer, 1987). Their 
learning behavior consists of activities carried out by team 
members through which a team obtains and processes data 
that allow it to adapt and improve. Examples of learning be- 
havior include seeking feedback, sharing information, asking 
for help, talking about errors, and experimenting. It is 
through these activities that teams can detect changes in 
the environment, learn about customers' requirements, im- 
prove members' collective understanding of a situation, or 
discover unexpected consequences of their previous actions. " (Edmondson 1999)

"A set of structural features-consisting of a clear compelling 
team goal, an enabling team design (including context sup- 
port such as adequate resources, information, and rewards), 
along with team leader behaviors such as coaching and di- 
rection setting-have been shown to increase team effec- 
tiveness (Hackman, 1987)"

"These 
cases suggest an asymmetry, in which high-learning teams 
can confront and work with design and other constraints to 
improve their situation, while low-learning teams are far 
more likely to get stuck and be unable to alter their situation 
without intervention. " (Edmondson 1999)

In order to create new value and competitive advantage in rapidly changing and uncertain organisational environments, new managerial imperative is growing focusing on teams. Thus, supporting teams in their work and understanding the aspects of team learning is required. (Edmondson 1999) Recent studies has moved the focus from individual learning to team learning

Fast-pace environment requires organisations to enhance the ability of teams to learn and create environment where learning is safe and possible. Uncertainty affects all working life, change is faster and job security lower. Psychological safety for individuals at work can be increased through great teams and teamwork. Employees should be engouraged to ask questions, seek for help and tolerate mistakes and uncertainty. (Edmondson 1999)

As Edmondson pointed out, psychologicl safety as a means to promote team performance is increasingly relevant both in future work and research. 

Definition of team consists of a small group of people who have supplementary knowledge and abilities compared to each other, and who share a goal, targets and way of working and approach. Team also shares a feeling of being responsible for its performance. (Edmondson, 1999)

Team opportunities and full advantage of team effort can be threatened by demanding schedules, long-standing habits and unwarranted assumptions. In addition, in order to function, team needs a clear purpose and vision what makes it a team and why it exists. Teams get energy from significant performance challenges regardless of where they are in the organisation. Set of shared, demanding performance goals usually form a team, and personal chemistry or willingness to form a team may boost that, yet teamwork is essential for an actual team. Indeed, in order to receive great results teams should focus on performance regardless of the organizational hierarchy or what team does. Thus, team performance may exceed the results of what could be achieved if employees were acting alone as individuals without the team effort.  (Katzenbach)

Organizational leaders play a great role in establishing strong team performance culture. This can be achieved through addressing and demanding performance that meets the need of the customers, employees and shareholders. Teams should not be fostered by the sake of the team only, rather should leaders clearly state how the team performance affects to customers and through that clearer performance ethics and cultures are fostered. In addition, even though people tend to have great sense of individualism, it does not have to bias the teamwork performance, as real teams find ways to support individual strengths and performance for shared goal. Furthermore, in order to team function properly and efficiently, discipline across the team and organisation is needed, focusing again on performance. "Groups become teams through disciplined action." According to Katzenbach great team performance consists of continuos work of shaping a common purpose, agreeing on performance goals, defining a common working approach, developing high level complementary skills and being transparent on the results.  (Katzenbach) 

Team psychological safety refers to Amy Edmondson's concept of team members shared belief team being safe for interpersonal risk-taking. Together with team efficacy these have great affect on team performance and learning in an organisational team work. According to Edmondson, team performance and outcomes can be shaped through team structure and shared beliefs. For instance, employee's willingness to take interpersonal risks depends highly on the experience of team safety and person's beliefs how others will respond in ideas or situations involving uncertainty. Team psychological safety refers to interpersonal trust among team but beyond that mutual respect and caring.  (Endomndson 1999)

According to Edmondson (1999) learning behaviour consists of seeking feedback, sharing information, asking for help, talking about errors and experimenting. These factors were also recognised from the data as factors affecting experimentation. Thus, experimentation behaviour seems to relate to learning of an individual, teams and organisations and can be supported by supporting these factors. 

Studies have shown the essence of psychological safety in teams, defining how structural and interpersonal characteristics both have effects on learning and performance in teams. Psychological safety can be boosted for instance through structural factors such as context support and team leader coaching affecting behavioural and performance outcomes. Furthermore, climate of safety and supportivenesss encourages employees to seek for feedback and ask for help in addition to admit and reflect mistakes.  (Edmondson 1999) 

As Edmondson suggested, studies in real work teams are required, and this study aims to add on this aspect studying factoring affecting experimentation behaviour in real work context. 

"Innovation teams not only form the engine and heart of the innovation process, they are also essential for encouraging the organization to accept the innovation result. Innovation team members come from many different relevant departments or functions within the company. " (Buijs, 2007)

"The selection of the innovation team leader (usually by the top management of the company) should be the starting point of the forming stage. In the ideal situation, the newly appointed innovation leader can pick out his or her own volunteers. The leader should at least have the power to veto unwanted team members. Innovation is about having the right people in your team." (Buijs 2007) Tämä pointtina, että kokeilemisenkin pitäisi olla vapaaehtoista? Vai meneekö liian kauas

"At the level
of teams, some studies have demonstrated that team members? collective view of support from their
leader is associated with the team?s success in creative endeavors (e.g., Amabile and Conti, 1999; Amabile
et al., 1996)."

However, another point of view to ideation is that truly novel ideas raise from individuals, being the ultimate source of any new idea or solution to a problem (Amabile et al. 2004).

"Robinson and Stern (1997)  propose that employees' capabilities can be enhanced by stimulating co-workers, who are more likely to challenge each other's ideas in
constructive ways."

"The development and implementation of creative ideas oftentimes requires input and support from multiple individuals or groups (Mumford et al., 2002)."

"As such, the social influences of others are important. For instance, several job attitudes are socially constructed based on the social labeling of work by others (Nemeth and Staw, 1989). In addition, the opinions of others with regards to one's own work can influence how an employee perceives her job and organization (Salancik and Pfeffer, 1978)."

"Agrell and Gustafson (1994) proposed that participation by team members should help ensure that new ideas are not abandoned."
"While leaders can play a key role in modeling creative behaviors, it is important for them to realize that teammates and coworkers also can exert a strong influence. Here, managers can use human resource practices to foster a team environment that supports and encourages creativity." (Shalley and Gilson 2004) 

"Monge, Cozzens, and Contractor (1992) found that group communication was positively related to the generation of innovative ideas."
"Theorists have proposed that the communication of ideas and information along with contact with diverse others should lead to higher levels of creativity (Amabile 1996; "

"These research findings suggest that simply changing the structure of teams (i.e. combining representatives of diverse function and tenure) will not improve performance. The team must find a way to garner the positive process effects of diversity and to reduce the negative direct effects. At the team level, greater negotiation and conflict resolution skills may be necessary. At the organization level, the team may need to be protected from external political pressures and rewarded for team, rather than functional, outcomes." (Ancona and Caldwell 1992) 

"For instance, research has found that structures that promote open, ongoing contact with external others or information seeking from different or multiple sources were related to creativity (e.g., Ancona and Caldwell, 1992 and Dougherty and Hardy, 1996)."

"We need the outcasts to provide those other channels we are missing." (Stenberg et al. 1997)

Under some circumstances, according to Monge et al. (1992) group communication is likely to increase innovation. Thus, leaders should consider managing wide range of formal and informal meetings and facilitated discussions in order to create opportunities. Furthermore, innovation occurs over time and is a dynamic process. Leader should be sensitive in which pace more managerial impact is needes, and in which pace of the process more freedom and autonomy should be allowed for employees. 

Amy Edmondson (1999) "introduces the construct 
of team psychological safety-a shared belief held by 
members of a team that the team is safe for interper- 
sonal risk taking-and models the effects of team psy- 
chological safety and team efficacy together on learning 
and performance in organizational work teams. Results 
of a study of 51 work teams in a manufacturing com- 
pany, measuring antecedent, process, and outcome vari- 
ables, show that team psychological safety is associated 
with learning behavior, but team efficacy is not, when 
controlling for team psychological safety. As predicted, 
learning behavior mediates between team psychological 
safety and team performance. The results support an in- 
tegrative perspective in which both team structures, such 
as context support and team leader coaching, and shared 
beliefs shape team outcomes.' "

"Structural and interpersonal factors have been viewed in the 
literature as alternative explanations for team effectiveness. 
This study supported, instead, an integrative perspective, in 
which structural and interpersonal characteristics both influ- 
ence learning and performance in teams. In particular, the 
results showed that psychological safety is a mechanism 
that helps explain how previously studied structural factors, 
such as context support and team leader coaching, influence 
behavioral and performance outcomes. "(Edmondson 1999)

\subsection{Individual learning}
According to experiential learning theory, learning is "the process whereby knowledge is created through the transformation of experiences." (Kolb) Experiential learning process stands as a way describe the central process of human adaptation to the social and physical environment. (Kolb) Learning involves total concept of human being - feeling, thinking, perceiving and behaving. (Jung, 1923) Learning should be conceived as a holistic adaptation process providing bridges across life situations and underlaying the lifelong process of learning. (Kolb)

However, learning may not be the first thought in mind when considering an immediate reaction to the situation or problem at hand, rather is it considered as performance. Furthermore, long-term adaptations to our previous experiences and beliefs is mainly considered as developing, not learning. Thus, when talking about developing an individual or an organisation, the question highly concerns and is related to learning. (Kolb)

According to Kolb (..) experience plays a central role in learning, and learning is a process consisting of experience, perception, cognition and behaviour. Immediate experience forms a basis for reflection and observation, following assimilation to a theory from which new implications for action are deducted. In order to create new experiences, these implications serve as guides. 

This Lewinian approach emphasises here and now -concrete experiments in testing and validating abstract concepts. Thus, experience of an individual is a focal point of learning, and giving personal meaning to abstract concepts, which can be afterwards shared with others. Furthermore, receiving feedback is considered essential in this approach for learning, as it serves a continuous process for goal-oriented action following evaluation of that action. Feedback can thus boost effective, goal-oriented learning process. (Kolb)

Kolb presents Dewey's model of Learning, which adds to Lewinian model how through learning the impulses, feelings, and desires of concrete experience into higher-order purposeful action is transformed. Both Lewin and Dewey consider learning as a dialectic process integrating experiences and concepts, observations and actions. 

Piaget has formed in 1970 model of learning and cognitive development, which presents similar aspects to Lewin, Dewey and Kolb. Piaget also stated how learning is an interactive process between individual and environment. Learning is a mutual process between accommodation of concepts or schemas to experiences around us and assimilation of events and experiences into existing concepts and schemas. This intelligent adaptation, learning, results from the tension between accommodation and assimilation. Through this tension growth and higher-level cognitive functioning occurs. 

Kolb summarises above models. First of all, learning should not be conceived in terms of outcomes, but it is best conceived as a process. Ideas are not fixed and and immutable elements of thoughts, but can be formed and re-formed through experience,  

Bringing the experiential learning into education implications, all learning can be considered as relearning. Thus, all learning situations should take into account people arriving from all different experiential backgrounds to what they build their new experiences and knowledge on. This partly explains very likely resistance to new ideas, as when new information and experiences are in contradiction to old beliefs and experiences, new ideas and information is more difficult to adapt. In the education process learner's old beliefs and theories should be brought out, examined and tested, following integration of the new models and refined ideas into learner's belief systems. (Kolb)

According to Piaget individual learning and adaptation of new ideas occurs through integration or substitution, integration leading to stronger part of learner's conception of the world, whereas substitution requires real questioning of previous conceptions, and thus might take longer for the learner to adopt. 

According to Kolb, learning is a process filled with tension and conflict, and new knowledge, skills and attitudes are achieved through experiential learning, which consists of four modes and required abilities of learners: concrete experience abilities, reflective observation abilities, abstract conceptualisation abilities and active experimentation. First of all, individuals must openly involve themselves in new experiences, reflect and observe them from various perspectives, create concepts that can be integrated into more abstract theories as well as they need to be able to use these reflections and theories in active daily decision-making and problem-solving. Quite a task and ability list for individuals, and this is to show how learning requires various abilities from different perspectives and to choose which learning perspective to use in a situation at hand. Furthermore, Kolb divides the learning into two dimensions: "first of all representing the concrete experiences at one end and abstract conceptualisation at the other. " The other dimension includes active experimentation at one end and reflective observation at other. Thus, the level of learning depends on the way in which conflicts ... Kesken, kolb page 15. One mode dominating the conflict resolving process leads to learning dominated by that mode. However, the aim is to create a synthesis between the different modes in order to boost the process of learning and personal development. 

Also Sydänmaanlakka emphasises different learning styles and modes, which should be understood and used variously in learning and problem-solving situations. 

In addition the meaning of environment in learning should be emphasised. Learning concerns of transaction between an individual and the environment, learning does not happen only inside of individual's thoughts, experiences and processes but is dependent on the real world environment. 

"What is of note here is that creative people consistently seem to express a pattern of characteristics that supports the confident exploration of alternatives under conditions of ambiguity (Mumford and Gustafson, 1988)."

According to Shalley and Gilson (2004), creative traits of individuals include "broad interests, independence of judgment, autonomy, and a firm sense of self as creative." In addition, Amabile (1988) has stated relevant skills required for individual creativity such as as the ability to think creatively, generate alternatives, engage in divergent thinking, or suspend judgment.

Shalley and Gilson (2004) argue that through developing extensive set of skills, employees learn to be more comfortable and confident in thinking from different perspectives, finding various alternative solutions, trying out novel things and seizing opportunities. 

According to Tierney et al. (1999), when employees enjoy creative tasks at hand, their level of creativity is high. Intrinsic motivation have been widely recognised as an important factor for creativity and innovation. (Tierney et al. 1999; Amabile 1988; Csikszentmihalyi 1999)

\section{Innovation to the rescue}

"Drawing on a multiyear research project and many case studies, the author analyzes the managerial practices of successful large companies and outlines some common patterns in their approach to technological innovation. These big companies, like many successful small entrepreneurs, accept the essential chaos of development. They pay close attention to their users' needs and desires, avoid detailed early technical or marketing plans, and allow entrepreneurial teams to pursue competing alternatives within a clearly conceived framework of goals and limits." (Quinn 1985)

"Clearly, a single enterprise cannot by itself develop or produce all the parts needed by such large new systems. And communications among the various groups making design and production decisions on components are always incomplete. The probability of error increases exponentially with complexity, while the system innovator's control over decisions decreases significantly-further escalating potential error costs and risks. Such forces inhibit innovation in large organizations. But proper management can lessen these effects.
--

"Motivate people to contribute ideas by 
making it safe to fail. Stress that the goal is to 
experiment constantly, fail early and often?
and learn as much as possible in the process. 
Convince people that they won?t be punished 
or humiliated if they speak up or make 
mistakes." (Amabile and Khaire 2008)

Many studies suggest that effective technological innovation develops hand-in-hand with customer demand" (Quinn 1985)

"The term social innovation, as used here, refers to
the generation and implementation of new ideas about
how people should organize interpersonal activities,
or social interactions, to meet one or more common
goals. As with other forms of innovation, the products
resulting from social innovation may vary with regard
to their breadth and impact (Mumford and Gustafson,
1988)." (Mumford 2002)

"With regard to idea generation, it appears that
a climate allowing autonomy and providing a mix of
intrinsic and extrinsic rewards while encouraging an
intimate involvement with the work itself contributes
to innovation (Amabile, 1997)"

Characteristics associated with innovation are integration of work units, decentralisation of control and professionalisation are likely to effect innovation in a way that through these suitable environment for innovation, dynamic idea exchange and implementation is created (Mumford 2002). 

Mumford (2002) presents four factors affecting social innovation: active exchange of ideas and information in supportive climate, ideas that are tangible, low-cost and can be at least guessed to be beneficial, support from upper level management, and lastly effective communication through the innovation process in order to make it from the idea to implementation.

"Of course, in a complex social situation, where many causes operate, not all of which are controllable (Katz and Kahn, 1978, tarkista onko tää), innovation may ultimately depend on boiling things down to their essential and focusing on those essential elements of the situation you can do something about. (Mumford 2002) Voisko tämä toimia pointtina sille, että kokeileminen auttaa keskittymään sellaiseen, ja vain sellaiseen, johon voi vaikuttaa itse. Eli discussioniin 
Mumford (2002) also argues in his study of Ben Franklin's social innovations, that one key factor in success was fast demonstrating, as was also called experimenting. JEE. 
"Virtue of the need for persuasive demonstrations, opportunism and showmanship may be essential personal requirements for driving social innovations" (Mumford 2002)

Innovation and creativity should not be confused to be the same: creativity is needed as a first step in creating something novel whereas innovation refers to the "implementation of those novel ideas in individual, team or organisational level (Shalley and Gilson 2004; Amabile 1996; Mumford and Gustafson 1988). 

According to Amabile (1988), individual creativity stands for an essential building block for organisational innovation. 

"Reviews the contributions to the understanding of creativity and innovation in organisations and interprets the implications for training and development. Highlights key and recurrent principles: the benefits of an integrated organisational approach, the right climate for creativity, appropriate incentives for innovators, a structured means of search and a systematic way to convert an opportunity into an innovation. It implies a broad range of skills development including: team working, communications, coaching, project management, learning to learn, visioning, change management and leadership. Even though techniques for the development of innovation and enhancing creativity in individuals are well founded, there are relatively few reports on the practice of mainstreaming creativity in an organisational setting. Likewise, although the transition from idea to innovation can be systematised, problems arise from customised applications that involve the management of the change process. " (Roffe 1999)

 innovativeness is positively related to the strength of superordinate identity in the team, encouragement to take risk, customers' influence, and active monitoring of the project by senior management. Beyond a moderate level, social cohesion among team members has a negative effect on innovativeness. The effect of superordinate identity on innovativeness is strengthened by encouragement to take risk and weakened by social cohesion. Functional diversity has no effect on innovativeness. (Sethi et al. 2001)
 
 "Innovativeness presumes a degree of creativity in the new product ideation and design processes." (Sethi et al. 2001)


\subsection{Creativity, intrinsic motivation and everyday problem solving}

"Psychometric instruments have documented the absence of correlations, above a threshold, between measured intelligence and creativity; computer simulations and ?on?line? studies of human subjects indicate that creative problem solving draws on the same information?processing skills as everyday problem solving; case studies reveal that creative individ?uals pursue a network of interrelated enterprises over the course of their productive lives and that most creative insights emerge gradually. Creative individuals exemplify a predictable set of personality traits and are motivated chiefly by intrinsic factors. Sociological and historiometric approaches document the conditions under which creative products are likely to emerge within a given society, and neurobiological perspectives suggest that certain neural structures and processes may be particularly associated with creative out?comes. A science of creativity is most likely to emerge from a synthesis of these different disciplinary perspectives." (Gardner 1988) Furthermore, ability to find problems has been related to generation of novel, alternative solutions (Runco and Okuda 1988), and problem finding skill has been indicated to be one of the best predictors of creativity in 'real world' activities, when studied 91 elementary school students (Okuda, Runco and Berger 1990). These findings suggest leaders, in order to enhance creativity of employees, to support and facilitate problem-construction (Redmond et al. 1993)

According to Isaksen (1983), in order to support employee's creativity, leaders should focus on creating and maintaining and environment of supportive empathy, respect, warmth, concreteness, genuiness, trust and flexibility. These factors have been combined to general and task-specific efficacy needs (Mumford and Gustafson 1988). Furthermore, through providing enough processing time for creating novel solutions is likely to enhance creative behaviour of employees (Isaksen 1983). As creativity refers to finding novel solutions and generating understanding of problems at hand, leaders could facilitate the process of resource allocation, feedback and task management in order to support employee's creative process (Mumford and Gustafson 1988). 

For students of creativity, there is no surprise in attaching self-efficacy to creative actions (Mumford and Gustafson 1988), yet recently problem construction processes has been recognised and combined to everyday problem-solving and real-world creativity (Getzel and Csikszentmihalyi, 1975; Runco and Okuda 1988). In their study Redmond et al. (1993) find how leaders supporting employees problem-finding and problem construction led to more unique and novel solutions. Leaders encouraged employees to find alternative solutions, approach problems from different perspectives and overall supporting several alternative problem-solving strategies (Redmond et al. 1993). In addition, study showed how through motivational mechanisms, such as self-set goals, involvement and commitment, problem construction may have positive influence on solution quality and originality. Thus, problem construction is likely to have its greatest impacts on performance when in the process employee is allowed to express his values, needs and interests (Redmond et al. 1993). 

Yet, leader alone is not able to boost creative solutions in employees: it is also a matter of personal characteristics, previous knowledge of the problem at hand and expertise in the field (Mumford and Gustafson 1988, Redmond et al. 1993). Thus, in order to achieve novel solutions and fresh ideas, leaders may seek employees who have great knowledge and expertise of problem at hand or provide employees education and possibilities to develop their problem construction skills (Redmond et al. 1993) and furthermore encourage approaching problems from various perspectives. 

Furthermore, supporting employee's feeling of self-efficacy is likely to improve creative skills of an employee (Redmond et al. 1993), and can be done through giving positive and realistic feedback, allowing adequate resources and physical support, clarifying task assignments, providing development support for employees, and assigning employees to appropriate tasks (Hennessey and Amabile 1988). However, often acknowledging employee's skills, potential and accomplishments is likely to push an employee to the track of creativity (Redmond et al. 1993). 

Furthermore, Redmond et al. (1993) argue that the understanding of individual's creativity and ways to shape it gives us guidelines when creating an environment and leadership that support organisational innovation tai support creativity and innovation in organisational settings. (TÄMÄ TÄRKEÄ YHTEYS luovuuden, innovation ja organisaation välillä!) 

"In a series of causal analyses, it was
found that divergent thinking exerted unique effects on
creative problem solving that could not be attributed
to intelligence or expertise. Intelligence and expertise,
however, also contributed to creative problem solving" (Vincent et al. 2002)

Divergent thinking refers to ability to come up with multiple alternative solutions to a problem at hand. (Guildord 1950) Guilford (1967) has argued that this divergent thinking is a key capacity affecting creative thinking"

"Expertise, specifically in the domain of organisational leadership also had a moderate direct effect on the expression of divergent thinking skills." (Vincent et al. 2002)

Kasof (1997) argued in his study that breadth of attention affects on creative performance of an individual. By breadth of attention Kasof refers to "number and range of stimuli attended to at any time." When breadth of attention being narrow, individuals are able to focus on narrow range of stimuli and are better at filtering redundant stimuli from awareness. However, those individuals with wide breadth of attention tend to be more aware of irrelevant or extraneous stimuli, these individuals are strongly affected by their environment and are highly arousable. 

Studies have shown how "creative ability is related to chronically wide spread of attention" (Kasof 1997)

"Inherently, the level of creativity required may be dependent on the job in question. For example, when examining the tasks performed by RetD professionals, major breakthroughs may be desirable and necessary. In contrast, for the jobs of cashiers or assembly line workers, an incremental change in how the work is done may be a desirable creative outcome." (Shalley and Gilson 2004)


"the success of creativity depends on the capabilities, pressures, resources, and sociotechnical system in which employees find themselves (Csikszentmihalyi, 1999). In order for creativity to occur, leadership needs to play an active role in fostering, encouraging, and supporting creativity. Hence, the role of leaders is to ensure that the structure of the work environment, the climate and culture, and the human resource practices (e.g., rewards, resources, goals, and expected evaluations) are such that creative outcomes can and do occur  (Shalley and Gilson 2004; Oldham and Cummings 1996; Mumford et al. 2002)

 "That is, it would be difficult to be creative in an area without having some experience and knowledge about what has historically been constituted as routine or the status quo. Although, in some cases, task familiarity could lead to more ?habitual? performance (Ford, 1996), it also can provide the needed opportunity to prepare for creativity through deliberate practice of task domain skills and activities." (Shalley and Gilson 2004) 
 
"Essentially, if creativity is a valued outcome and employees believe this to be true, they should be more willing to experiment with new ideas, more open to communicating and seeking input from others about new ideas, and overall behave in ways that will lead to creative outcomes." (Shalley and Gilson 2004) 

Organizational environment plays a major role in employees' creative skills, and such stifling factors may be positive challenge at work, encouragement from organisational level, support from work group as well as supervisory encouragement. (Amabile 1997) Furthermore, organisational impediments can lead to decreased level of creativity. 

Amabile and Khaire (2008) argue managing creativity actually being managing for creativity rather than managing creativity. .

\section{Experimentation as a method for innovation and learning}

Normal business is considered as repetition, risk-avoidance and focusing on business outcomes (Buijs, 2007), while innovation requires novel solutions, thinking out of the box, risk-taking, breaking the rules, challenging the status quo and questioning the future (Burns and Stalker 1961; Kanter 1984, March 1991)

According to Buijs (2007), innovation consists of coming up with novel ideas and implementing them. Ideating begins with exploring, developing and implementing the ideas, following introducing the ideas, which have turned into products or services, into the marketplace. 

Innovation process can be approached from several angles: first of all, content of the innovation has to be clear - whether the purpose is to innovate new products, manufacturing processes, ways of organising or ways of dealing with people. Secondly, psychological process of the innovation team has to be understood, essential being shared understanding, level of comfort with ambiguity and degree of trust between team members. Thirdly, creative process of the team, meaning idea producing process, needs to be understood and efficiently facilitated. Finally, the role of leading plays a major role, and together with playful attitude innovation process is likely to succeed. (Buijs, 2007)

"Usually, the innovation process is described as a series of stages through which an idea is processed. During each stage, certain activities are executed in order to improve the quality of the idea and to let the idea grow. Each stage ends with a gate, in which the idea is checked and evaluated. It is then decided if the idea can be carried over into the next stage, if the idea needs to be further developed in the current stage, or if the idea needs to be abandoned altogether.---
Although stage-gate models are describing the process an idea is going through, the evaluation steps are useful not only for evaluating the quality of the ideas, but also for checking and evaluating the process itself. What went well? What can be improved? And what has been learned? Reflection occurs during evaluation and learning occurs during reflection. After all, every innovation process is an organizational learning process " (Buijs, 2007)

Several factors have been recognized to affect on organisational innovation, yet many researchers have stated leadership behaviour being one of the most important. (Dong et al. 2003; Amabile, 1998; Jung, 2001; Mumford and Gustafson 1988) Dong et al. (2003) identify four hypotheses how top managers' leadership styles may affect both directly and indirectly their companies' ability to innovate. Indirectly here stands for instance for leader's possibility to empower employees and build organisational climate optimal for innovation. The study shows a positive relation between transformational leadership style and empowerment as well as innovation-supporting organisational climate. 


As in todays' business world it has been widely recognised how creativity and innovation are essential for business growth, researchers have studied factors affecting creativity and innovation in organisations. Amabile (1998) has identified three factors as being important for stimulating creative behaviour in individuals and organisations: individuals' intellectual capacity (creative thinking skills), expertise based on past experience and creativity-supporting work environment. Furthermore, Oldham and Cummings (1996) consider creativity skills and characteristics of and individual as important, yet they add the importance of characteristics of organisational context such as job complexity, supportive supervision or controlling supervision. 


Experimentation calls for creativity. 
"When individuals access a variety of alternatives, example solutions, or potentially related ideas, they are more likely to make connections that lead them to be creative (Amabile et al. 1996)"

"In addition, skills such as problem finding, problem construction, combination, and idea evaluation are important for creativity (Mumford et al. 2002, Vincent et al., 2002).

"In contrast to personality traits and creativity relevant skills, domain-specific knowledge reflects an individual's level of education, training, experience, and knowledge within a particular context (Gardner, 1993). "

Several factors form the basis of creativity skills of an individual, such as personality, technical knowledge, expertise, motives, and the supervisor's feedback style. In group level factors form of task structure, communication styles and task autonomy, and finally in organisational level strategy, structure, culture, climate and available resources all affect how creative actions are encountered. (Jung et al. 2003) 

Creativity is not restricted to artistic occupations only; it is required in various professions in which tasks presented involve complex, ill-designed problems where novel solutions are needed and status quo challenged (Mumford and Gustafson 1988). 

According to Mumford (2002), actually, idea implementation may require even more creativity than idea generation and Vincent et al (2002) states the two key sets of processes that are involved in creative work: : "(a) creative processes or the activities underlying initial idea generation and (b) innovation processes or the activities underlying the implementation of new ideas"

"Implementation, of course, depends on the evaluation of ideas (Runco 1994)."

Amabile (1996,1998) defines creative thinking as a way how people approach problems at hand and come up with solutions. Creative thinking does not stand for intellectual capacity of an individual to create something new but rather as a combination of past experiences which creates expertise and the ability to apply creative thinking skills to these experiences and invent new solutions. 

Yet, intrinsic motivation of an individual has been recognised one of the most essential factors increasing creativity 

When dealing with novel solutions and challenging status quo, we are dealing with innovations. In order for company and its employees to be innovative, they need to take risks. Yet, at the same time usual management processes avoid risk-taking and focus on managing daily routine business. As Quin (1985) stated it in his Harvard Business Review article: we love innovation and we urge for innovation, but we can tolerate it only if it is controllable and results everything remaining the same. (Quin 1985). 

"Talking about innovation as a single process is misleading. Innovation is a multi-faceted process that is full of contradictions. It is simultaneously hard and soft, nice and nasty, fun and serious. Innovation involves technology and people, marketing thinking about manufacturing and manufacturing thinking about marketing" (Buijs 2007)

"Within the broad framework I have described, major innovations are best managed as incremental, goal-oriented, interactive learning processes." (Quinn 1985)

Anriopoulos and Lowe (2000) define in their study for elements of perpetual challenging - a way to enhance creativity and innovation in an organisation. First of the elements is adventuring, which refers to "the process through which individuals are encouraged to explore uncertainty, so that they can generate innovative solutions. Adventuring is the only process, which occurs implicitly when employees' goal is the generation of new ideas." Furthermore, they define three categories of adventuring: introspecting, scenario making and experimenting. 
"Creative employees,
who aim to do original work,need tohave the
basic knowledge of their specific field. In
other words, they need to know what has
already been discovered, so they will be able
to go beyond that point." Jatkaa pointtia, että perustietämyksen olisi hyvä olla kunnossa työntekijän ideoinnissa. 

"Scenario making refers to the development of
possible routes to tackle a particular
situation." In addition, scenario making attempts to clear what is already known about problem or situation at hand, and especially what is not known. --
Experimenting
occurs when employees utilise their full
potential by being involved in a personal
trial and error process. It refers to the
employees' ability to test the different
scenarios generated by different concepts in
terms of images or ideas, which can fit to the
parameters set by the client or the industry
within which their company operates, and
eventually decide the most appropriate one.--
the research has
shown that incremental risk is very often
cherished by creative organisations because
it stretches employees' capabilities and
consequently provides the basis upon which
employees can develop new knowledge and
skills to be used in other projects. By
stretching their creative capabilities creative
organisations try to enhance the quality of
services offered to their clients every time an
opportunity arises. --
Nevertheless, creative
employees need to know that they operate in
an environment where failure is not
penalised, as long as the creative process and
analytical methods are effectively used. This
process is labelled as safety netting since
from the interviews it was apparent that
creative employees should be in an
environment which tolerates failure. --
Conceptual confronting refers to the
employees' ability to question each other's
ideas so that their full potentials are utilised.
Conceptual confronting can be internal,
which refers to the questioning of each
other's ideas within the company, and
external, which occurs when informal and
casual debates with people from other
disciplines take place--
This research has revealed that a pre-set
obstacle like a deadline can stimulate
creativity, as employees' focus can be
narrowed to the issue in question, which
needs to be solved immediately.(Andriopoulos and Lowe 2000)

\section{Factors hindering experimentation and learning}

Several factors may affect on the gap between idea and action in employees of an organisation. 

One essential factor is the beliefs, emotions and actions of an employee. For instance, when feeling under risk or pressure, an employee is likely to inhibit learning behaviour as a result of feeling the fear of being rejected or feeling they are placing themselves at risk. This may occur in a situation where an employee should ask for help, yet is afraid of admitting he lacks abilities, skills or knowledge. (Endomndson 1999). In addition, admitting mistakes, asking for help and seeking feedback are all relevant abilities in the recent organisational world, yet threatening for an individual's image of himself and his skills (Brown 1990). Indeed,  learning is often inhibited as people tend to act in a way that inhibits learning when facing the potential for embarrassment of threat, even though their transparency and honesty would be highly important for the behaviour of the team. (Argyris 1982)

The phenomenon of threat of employees in organisations is widely studied and consensus is rising how threat effects on cognitive and behavioural flexibility and responsibility in reducing manner. 

"Top management isolation. Many senior executives in big companies have little contact with conditions on the factory floor or with customers who might influence their thinking about technological innovation. Since risk perception is inversely related to familiarity and experience, financially oriented top managers are likely to perceive technological innovations as more problematic than acquisitions that may be just as risky but that will appear more familiar." (Hayes and Garvin 1982)

"Intolerance of fanatics. Big companies often view entrepreneurial fanatics as embarrassments or troublemakers. Many major cities are now ringed by companies founded by these "nonteam" players-often to the regret of their former employers.
Short time horizons. The perceived corporate need to report a continuous stream of quarterly profits conflicts with the long time spans that major innovations normally require. Such pressures often make publicly owned companies favor quick marketing fixes, cost cutting, and acquisition strategies over process, product, or quality innovations that would yield much more in the long run.
Accounting practices. By assessing all its direct, indirect, overhead, overtime, and service costs against a project, large corporations have much higher development expenses compared with entrepreneurs working in garages. A project in a big company can quickly become an exposed political target, its potential net present value may sink unacceptably, and an entry into small markets may not justify its sunk costs. An otherwise viable project may soon founder and disappear.
Excessive rationalism. Managers in big companies often seek orderly advance through early market research studies or PERT planning. Rather than managing the inevitable chaos of innovation productively, these managers soon drive out the very things that lead to innovation in order to prove their announced plans.
Excessive bureaucracy. In the name of efficiency, bureaucratic structures require many approvals and cause delays at every turn. Experiments that a small company can perform in hours may take days or weeks in large organizations. The interactive feedback that fosters innovation is lost, important time windows can be missed, and real costs and risks rise for the corporation.
Inappropriate incentives. Reward and control systems in most big companies are designed to minimize surprises. Yet innovation, by definition, is full of surprises. It often disrupts well-laid plans, accepted power patterns, and entrenched organizational behavior at high costs to many. Few large companies make millionaires of those who create such disruptions, however profitable the innovations may turn out to be. When control systems neither penalize opportunities missed nor reward risks taken, the results are predictable." (Quinn 1982 on barriers to innovation) 

"Expertise,
intelligence, and divergent thinking did not exert
direct effects on performance. Instead, the effects on
performance were mediated by problem-solving
activities within the domain under consideration?
specifically, idea generation and idea implementation.
These observations about the role of expertise,
intelligence, and divergent thinking point to a broader
conclusion. The generation and implementation of
solutions to real-world problems of the sort examined
in this study, like performance on most significant
occupational tasks, are a complex phenomenon.
Both divergent thinking and expertise influence leader performance but only through the leader's ability to generate and implement solutions to significant organisational problems." (Vincent et al. 2002)

\section{Factors supporting experimentation and learning}

Creativity and innovation have gained wider acceptance as important factors creating value in organisational performance (Mumford et al. 2002). Creativity and innovation have for instance been studied to have enhancing impact to organisations profit and growth (Nystrom 1990). 

According to Garvin et al. (2008) in learning organisation employees excel at creating, acquiring and transferring knowledge. They define building blocks for learning organisation: supportive learning environment, concrete learning processes and practices and leadership behaviour that reinforces learning. Building blocks can be considered and measured as independent components yet each of them vital to the whole. In order to improve long-term learning of an organisation, strengths and weaknesses of an organisation and its unit needs to be recognised. 

Likewise, Mumford and Gustafson (1988) emphasise the meaning of environmental variables as a means to support employee's creativity by providing resources to stimulate fresh ideas of employees. Furthermore, strong positive relations between organisational environmental variables have been found; organisational encouragement as well as support for innovation and creativity from team improve employee's creativity (Amabile et al. 1996) 

In order experimenting to happen in an organisation safe and supportive environment has to be created. According to Garvin et al. (2008) supportive learning environment consists of four characteristics: psychological safety, appreciation of differences, openness to new ideas and time for reflection. 

Futhermore, Amabile (1998) suggests changes in organisational environment are likely to boost intrinsic motivation of an employee leading to increased creativity skills. Role of leaders and managers is essential; being a key person in organising group work and processes a leader may encourage employees to achieve shared goals. 

"Atmosphere and vision. Continuous innovation occurs largely because top executives appreciate innovation and manage their company's value system and atmosphere to support it. 
Because familiarity can foster understanding and psychological comfort, engineering and scientific leaders are often those who create atmospheres supportive of innovation, especially in a company's early life. Executive vision is more important than a particular management background.
Small, flat organizations. The most innovative large companies in my sample try to keep the total organization flat and project teams small. Development teams normally include only six or seven key people. This number seems to constitute a critical mass of skills while fostering maximum communication and commitment among members
Since it takes a chain of yesses and only one no to kill a project, jeopardy multiplies as management layers increase.
Recognizing the inadequacies of theory, innovative enterprises seem to move faster from paper studies to physical testing than do noninnovative enterprises. When possible, they encourage several prototype programs to proceed in parallel. Sony pursued 10 major options in developing its videotape recorder technology. Each option had two to three subsystem alternatives. Such redundancy helps the company cope with uncertainties in development, motivates people through competition, and improves the amount and quality of information available for making final choices on scale-ups or introductions.
Many experienced big companies are relying less on early market research and more on interactive development with lead customers. Hewlett-Packard, 3M, Sony, and Raychem frequently introduce radically new products through small teams that work closely with lead customers. These teams learn from their customers' needs and innovations, and rapidly modify designs and entry strategies based on this information.
Formal market analyses continue to be useful for extending product lines, but they are often misleading when applied to radical innovations. 
To allocate resources for innovation strategically, managers need to define the broad, long-term actions within and across divisions necessary to achieve their visions. They should determine which positions to hold at all costs, where to fall back, and where to expand initially and in the more distant future.
A company's strategy may often require investing most resources in current lines. But sufficient resources should also be invested in patterns that ensure intermediate and long-term growth; provide defenses against possible government, labor, competitive, or activist challenges; and generate needed organizational, technical, and external relations flexibilities to handle unforeseen opportunities or threats. Sophisticated portfolio planning within and among divisions can protect both current returns and future prospects-the two critical bases for that most cherished goal, high price-earnings ratios." (Quinn 1985)

Mumford and Gustafson (1988) have studied the gap between an idea an action, and came to conclusion it depending on various attributes relating to individual and circumstances. 

Psychological safety means that learning of employees occurs when employees do not fear being rejected, ask naive questions, make mistakes or present viewpoint of minority. Psychologically safe environment enables employees comfortably to express their thoughts at work. Appreciation of differences is important as opening minds for different ideas and world views increases both energy and motivation, brings out fresh thinking. Learning occurs when employees become aware of opposing ideas in a safe environment, and additionally openness to new ideas is required. Novel approaches are relevant for learning, thus employees should be encouraged in risk-taking and exploring and testing uncertain things. Lastly, through providing time for reflection learning in safe environment occurs. Instead of looking and judging by numbers of hours of work or results employees should be given enough time to reflect their work. Analytic and creative thinking will not occur under stress, heavy workload and too tight schedule. Under stress ability to recognise and react to problems and learn from experiences deteriorates. In supportive learning environment time for reflection is allowed. Garvin et al. (2008)

Similar characteristics were found in data and are further presented in chapter xx. 

Edmondson in her research (1999) studied whether beliefs in interpersonal context varied between the same organisation and whether it affected in team outcomes

According to Garvin et al. (2008) second building block of organisational learning, consists of concrete learning processes and practices. It includes experimentation, information collection, analysis, education and training and information transfer. Organizational learning can be supported trough concrete steps and activities which are tested and further developed through experimentations. Furthermore, information and intelligence about customers as well as technological trends should be collected systematically and further analysed focusing on identifying problems and solving them. Training and education of new and established employees is an essential part of practices and processes. Finally, through transparent and meaningful knowledge sharing organisational learning can be enhanced, focus being on clear, well-defined and working communication systems that employees can easily relate and feel useful. Concrete processes together with efficient knowledge sharing methods ensures that essential information is available quickly and efficiently for employees who need it. 

 Thus, emphasis should be put on creating and defining concrete learning processes and practices.
 
Thirdly, leadership behaviour should reinforce learning. Behavior of leaders is highly related to the performance of employees (Kim and Mauborgne 2014) and organisational learning (Garvin et al. 2008). In order to encourage employees to learn, leaders should prompt dialogue and debate, ask questions and listen to employees. Example of the leaders was also recocgized from the data as an important factor affecting experimentation behaviour, and Garvin et al. also emphasises how through own example leaders can encourage employees to offer new ideas and options. 

These three building blocks overlap to some degree and reinforce one another. For instance, leadership behaviour helps in creating supportive learning environment and this supports managers and employees in creating and defining concrete learning processes and practices. Furthermore, concrete processes support leaders behaviour in a way that fosters learning and through own example cultivates that behaviour to others. 

Supportive leadership behaviour alone is not sufficient for guarantee organisational learning. Garvin et al. (2008) emphasise how organisations are not monolithic and managers should be sensitive to differences in culture, department and units. In addition to cultural differences, learning requires clear and targeted processes and practices. Furthermore, learning should be considered as multidimensional, thus organisational forces should not be solely focused on a single area but to consider presented building blocks as a whole.

Intrinsic motivation of an individual is one of the most powerful tools to creative action non-traditional thinking (Amabile 1996; Deci and Ryan 1985, Jung 2001). 

Leaders supporting new ideas and idea exchange has been related to enhancing creativity especially among those employees who showed disposition towards creativity (Oldham and Cummings 1996). 

Expertise 
- Expert performance and its affects on implementing ideas (Ericsson and Charness 1994)

Collins and Amabile (1999) 

Creative work is demanding and time-consuming (Mumford et al. 2002) as well as requires attention over long periods of time involving high level of ambiguity and stress (Kasof 1997)
Thus, creative work is resource intensive (Mumford et al. 2002) involving risk. 
Through setting goals and doing small prototypes uncertainty can be reduced (Mumford 2002). 

Suoraan Mumfordin et al (2002) artsusta seuraavat kolme kappaletta, kerää näistä olennaiset point it, kolme kohtaa!: "This integrative style seems to involve three crucial elements. The first major element of this style is idea generation. Indeed, the bulk of the available research on creative leadership stresses the role of the leader in facilitating others' idea generation. Thus, intellectual stimulation, or the application of creative problem solving techniques in guiding others, seems required along with support for new ideas, involvement with the people in developing ideas, and granting these people the freedom to pursue the ideas thus generated. Not only must leaders help people generate ideas, they must construct an environment where such ideas and likely to emerge. Thus, leaders must insure diversity in the group, open communications, and through role modeling, crisis management, and policy decisions seek to create a climate and culture where people are likely to generate and pursue new ideas.

These idea structuring activities, to contrast to idea generation, tend to be indirect involving the creation of action, or project, frameworks so as to maximize the autonomy of the individuals doing the work. In other words, it may be more useful to set a deadline than to show someone how to meet this deadline.

The third aspect of this style of leadership is idea promotion. Idea promotion involves gathering support form the broader organization for the creative enterprise as a whole as well as implementation of a specific idea or project. For the leaders of creative people, these promotional activities are essential primarily because they insure that the resources needed to carryout the work will be available"

According to Amabile et al. (2004) and their componential theory on creativity, "the support provided by immediate supervisors exerts
an influence on subordinates? creativity through direct help with the project, the development of
subordinate expertise, and the enhancement of subordinate intrinsic motivation. The componential
theory proposes that positive behaviors of supervisors include serving as a good work model, planning and setting goals appropriately, supporting the work group within the organization, communicating and
interacting well with the work group, valuing individual contributions to the project, providing
constructive feedback, showing confidence in the work group, and being open to new ideas. Thus, leader support behaviours should include both instrumental (or task-oriented) and socio-emotional (relationship-oriented actions." 

"Designing the work environment so that the natural flow of traffic through the building brings different functional areas in contact with each other could help facilitate and increase informal conversations." (Shalley and Gilson 2004) 

\subsection{Psychological safety}

"Team psychological safety should facilitate learning behavior 
in work teams because it alleviates excessive concern about 
others' reactions to actions that have the potential for em- 
barrassment or threat, which learning behaviors often have. 
For example, team members may be unwilling to bring up 
errors that could help the team make subsequent changes 
because they are concerned about being seen as incompe- 
tent, which allows them to ignore or discount the negative 
consequences of their silence for team performance. In con- 
trast, if they respect and feel respected by other team mem- 
bers and feel confident that team members will not hold the 
error against them, the benefits of speaking up are likely to 
be given more weight." (Edmondson 1999)

Interestingly, studies show how nominal groups perform remarkably better in ideation and brainstorming processes by producing greater amount of ideas than real groups. This may be due to the learnt practices and norms of a real work group, fear of failure that prevents free idea exchange and fear of evaluation and others judgement when suggesting creative solutions. (Jung 2001) 

Accordingly, Amabile (1998) suggested that creative thinking can be encouraged by shaping organisational culture such that employees feel encouraged to tell their ideas out loud freely and without judging, increasing idea exchange and discussion about them. 

"According to Mumford et al. (2002), organizational climate and culture represent collective social construction, over which leaders have substantial control and influence. Jung (2001) also views managers as playing key roles in developing, transforming, and institutionalizing organizational culture."

"Schein (1992) argues that as organizational founders and leaders communicate what they believe to be right and wrong, these personal beliefs become part of the organization's climate and culture. "

"In summary, there is substantial theoretical support for expecting that leaders play a major role in establishing an innovative organizational culture and facilitating creativity in organizations. In addition, the role of creativity and innovation in determining organizational performance has been well established. However, as Mumford et al. (2002) have observed, conspicuously absent from the literature are empirical studies on the link between leadership and innovation at the organizational level while incorporating contextual variables. Considering that researchers have emphasized the transformational leader's role in creating an intellectually stimulating work environment, it is surprising to find that no study has yet examined how a transformational leadership style affects innovative organizational climate and how it further affects creativity and organizational innovation. " (Jung et al. 2003)

"Previous research presents some intriguing evidence that people?s perceptions of the work
environment created by their team leaders and, in particular, their perceptions of instrumental and
socioemotional support, relate to their creativity" (Oldham and Cummings 1996)

According to Mumford and Gustafson (1988) "environmental variables can affect creativity by structuring problem solving efforts and by facilitating development and application of the basic generating processes giving rise to novel problem solutions." 
Furthermore, environmental factors can stimulate individuals' willingness or motivation to pursue new ideas and social environment is the source for resources and support needed in implementing novel ideas. (Mumford and Gustafson 1988)

"The data also suggest 
that team psychological safety is something beyond interper- 
sonal trust; there was evidence of a coherent interpersonal 
climate within each group characterized by the absence or 
presence of a blend of trust, respect for each other's com- 
petence, and caring about each other as people. But building 
trust may be an important ingredient in creating a climate of 
psychological safety. -- Although building trust may not nec- 
essarily create a climate of mutual respect and caring, trust 
may provide a foundation for further development of the in- 
terpersonal beliefs that constitute team psychological safety. " (Edmondson 1999)

"Effective team leader coaching and con- 
text support, such as access to information and resources, 
appear to contribute to, but not to fully shape, an environ- 
ment in which team members can develop shared beliefs 
that well-intentioned interpersonal risks will not be punished, 
and these beliefs enable team members to take proactive 
learning-oriented action, which in turn fosters effective per- 
formance. " (Edmonsdon 1999)

\subsection{Leadership behaviour}

Leaders have a great influence on employees on several levels and through several mechanisms, such as role modelling, goal definition, reward allocation, resource distribution, defining norms and values of the company, showing the way to interact as a group, condition employees' perceptions of work environment and being the lead decision maker on organisational structure and procedures (Avolio and Bass 1988) Tämä aika suoraan Amabile et al. 2004, joten muokkaa vähän! As a result of these, leaders have a great impact on employees' behaviour, thus leaders may have a significant effect on employees' creativity. (Amabile 1988) 

Leadership style has great impact on organisational innovation and creativity. For instance, leaders shape and define the goals and working context (Amabile 1998; Redmond 1993). Through a long-term vision (separated from short-term business outcomes, which usually focuses on quarterly profit), leader's are able to direct employee's efforts towards creativity and innovative work processes leading to likeminded outcomes (Amabile 1996)
Furthermore, in shaping and influencing organisational culture, leaders are key actors (Schein, 2010) for instance by sustaining and nurturing organisational climate that supports and encourages creative efforts and learning (Yukl, 2002). 

According to Redmond (1993) a leader can have an affect on employee's level of creativity through leadership behaviours such as problem construction, learning goals and feelings of self-efficacy. 

In addition, creating and supporting a reward-system that values creative performance, provides both intrinsic and extrinsic rewards for employee's efforts to learn new skills and to challenge status quo by experimenting new approaches, employees are constantly willing to engage in creative endeavours (Jung, 2001; Mumford and Gustafson 1988)

As Mumford et al (2002) put it, organisations may now need jazz group leaders rather than orchestra directors. 

Leaders can affect employees' creativity and innovation skills both directly and indirectly (Jung et al. 2003). By stimulating employee's intrinsic motivation and higher level needs leaders are able to affect directly on employees' creativity (Tierney et al. 1999), where indirect way may be through establishing a work environment where new ways of doing are encouraged and failure is not being punished (1996). 

Leadership behaviour is only recently recognized as essential part of enhancing creativity and innovation skills of employees (Mumford et al. 2002). Mumford et al (2002) suggests this may be due to our romantic perception of creative act, which defines creativity as an heroic act of an individual and leaders only being a hindrance to the creativity of an individual. Furthermore, other reason to this may be found in the conventional models of leadership, which not encourage employees to challenge the status quo but to achieve required goals. 

Conventional leadership behaviour focuses on internal activities within the team, whereas innovative team leader needs various set of skills and approaches in order to encourage developing and growing of teams and individuals. For instance, according to Barckzak and Wilemon (1989) leaders of innovative teams utilise wide range of familiar and unfamiliar techniques in order to accomplish the team objectives, whereas leaders of operating teams use only a few familiar techniques. Even though in this study innovation teams were not studied, similar elements of developing by experimenting and encouragement for that may be recognised, when dealing with new tasks and developing something which result is uncertain. 

Also Bujis (2007) states how leaders dealing with uncertain and new innovations should stay certain about uncertainties and provide a safe environment and encourage employees to work on current task comfortably. Thus, high level of tolerance for dealing with different states of minds and various personal feelings is required from a leader. 

Leaders and their behaviour have great influence on the creativity and innovation ability of employees (eg. Mumford et al. 2002; Jung 2001; Amabile 1998).

In order to encourage creativity and experimenting in teams, leaders should lead by example and act as role models. Leaders should consider their own behaviour and actions in a way that stimulates employees to new and innovative, creative approaches to problems. In addition, they can even request creative and innovative solutions form the team, which may lead to better results in creativity of individuals (Amabile et al. 2002). Lack of time for creativity and developing is not usually considered as part of work, thus time allocated for that is needed, as was also found in the empirical part of the thesis.   (Waldman, mum for, amiable, khaire)

Big ideas do not hatch overnight and creative thinking requires time. Leaders should allow team members time to think creatively, as according to studies under pressure creativity actually falls into decline even though individuals may feel more creative, yet actually they are perhaps working more and getting things done. According to this study, employees were clearly less creative while time pressure increased. (Amabile et al 2002) 

Furthremore, leaders can assist their employees by recognising times with high pressure, and allowing employees to focus on certain thing at a time, leaving the expectations of creativity and new ideas into the future moment, when time pressure has decreased. On the other hand, if creativity is required under stress, leader should transparently explain the importance and reasons behind the strict schedule and required goals. Thus an employee may relate to the problem at hand and engage better at his work. Indeed, helping people to understand the importance of work is essential especially under high time pressure. (Amabile et al. 2002)

Through encouraging employees in risk-taking and making mistakes, leaders are likely to boost innovation. (Farson and Keyes, 2002)

Failure as a part of innovation and development process begins to be generally recognised and approved (Farson and Keyes, 2002). Succeeding companies even thrive for failure in order to learn fast and find the best practices and business models. For instance, credit company Capital One conducts continually large amount of market experiments. They now most of the tests will not pay off, yet they also know how much can be learned about customers and markets from failed tests in early phase of development.  (Farson and Keyes, 2002). Yet leaders fail in showing their employees the support and tools for failing fast and early enough. Failing in a personal matter remains a difficult subject, as failing never feels exceptionally great, and often employees still consider failed work as failing personally. (Farson and Keyes, 2002) 

Failure-tolerant leaders put effort on explaining to employees how important part failure is to the development process as a whole, and how failing actually refers to a point where surprising, failed outcomes are not reflected and further analysed in order to learn. Permorming accordingly, admitting own failures and not chasing anyone to blame, failure-tolerant leaders encourage failure, lower the threshold and ease the fear of failing of employees. (Farson and Keyes, 2002)

Naturally management need to take seriously issues about safety and health, yet most of the failures should be seen as opportunities for growth. Furthermore, failure-tolerant leaders treat success and failure similarly, analysing and reflecting the outcomes in order to grow the intellectual capital of the team, including experience, knowledge and creativity. Other characteristics of failure-tolerant leaders are being rather collaborative than controlling, listening carefully, seeing the bigger picture, asking questions and focusing on the development and future rather than blaming on mistakes. In addition, in order to gain empathy and trust among employees, leader should admit their own mistakes, as it shows self-confidence and honesty, assisting in forming closer ties with employees. Vulnerability and transparency play a major role in trustworthy relationship between leader and employees. (Farson and Keyes, 2002)

Through the green light given and their own example leaders can change the focus from success and failure into thinking in terms of learning and experience. (Farson and Keyes, 2002)

As Amabile and Khaire (2008) draw a poetical picture how leader cannot manage creativity, but manages for creativity. Furthermore, they suggest that culture that fosters creativity includes leadership that enables collaboration, enhances diversity, encourages ideation, maps the stages of creativity to different needs, accepts inability and utility of failure and motivates employees with intellectual challenges. 

In experimentation process, employees need to contribute imagination, and this may require new kind of encouragement for creativity from the leaders. (Amabile and Khaire, 2008) Much success rises from employee's own initiatives, which results from wide amount of autonomy at work. 

Change from authority-based leadership to collaboration with employees has occurred in literature and in practice (Amabile and Khaire, 2008; Farson and Keyes, 2002). 

A culture of creativity can be fostered in an organisation through opening the organisation to diverse perspectives and openness to various ideas. This calls for safe environment for employees to share their thinking from different fields of expertise. Furthermore, engouraging passion and knowledge of an employee is likely to result in more creative action at work,  (Amabile and Khaire, 2008)

Allow experimenting 

Empowering employees is an essential tasks of leaders, through which a work environment is created where employees desire to seek innovative approaches to per form their work tasks (Jung et al. 2003). Transformational leaders encourage employees to participate in developing by highlighting the importance of cooperation, providing the opportunity to learn from shared experience and allowing employees to perform necessary actions in order to be more effective (Bass, 1990). Furthermore, autonomy and freedom to perform essential tasks has major effects on organisational creativity, as individuals are more likely to produce creative work when having the feeling of personal control over how to approach given tasks (Amabile et al. 1996) 
Yet, in order to maintain organisational innovation and risk-taking, autonomy given to an employee can not be in contradiction with fear of failure or discouragement towards challenging status quo or trying out novel solutions (Yukl, 2002). Thus, organisational climate has to support and encourage innovation (Mumford and Gustafson 1988) by valuing initiative and innovative approaches that support employees in risk-taking, accepting challenging assignments and stimulate intrinsic motivation towards work (Jung et al. 2003).

Provide clear focus (Barckzak and Wilemon 1989)

Transformational leadership refers to leadership style and processes which emphasises longer-term and vision-based motivational processes (Bass and Avolio 1997). Furthermore, through offering an explanation of the importance and value of the work, leaders encourage their employees' to think beyond self-interest (Yukl, 2002) 

Jung (2001) has studied how leadership style affects group's creativity and performance by comparing transactional and transformational leadership styles. Transformational leader refers to a leader who encourages divergent thinking and looking at problems from unconventional perspectives, while providing and explaining clearly defined goals and facilitating the innovation process of employees (Bass and Avolio 1990). Furthermore, development of clear long-term vision and practises supporting the way to achieve it is essential characteristic of transformational leaders (Avolio and Bass 1988). The relationship between transformational leader and an employee is active and emotionally attached (Avolio and Bass 1988) and through the strong attachment resulting from tight relationship leaders can better support employees in using their personal values and self-concepts in the way that employees can pursue higher level performance and fulfil personal needs through the work. This focus of transformational leadership on value alignment is likely to lead to the root of intrinsic motivation of an employee (Gardner and Avolio 1998), which is considered as one of the key elements in creative thinking and innovation skills of an employee (Jung 2001; Amabile 1998, Deci and Ryan 1985)

Transformational leaders can build environments that support creative actions (Sosik and al 1998; Avolio and Bass 1988). According to Sosik and al (1998) one key characteristic of transformational leader is the intellectual stimulation, which is likely to encourage creativity and divergent thinking leading to unconventional solutions to problems at hand. 

According to Bass and Avolio (1994) transformational leadership consists of four unique yet interrelated behavioural components: inspirational motivation (articulating long-term vision), intellectual stimulation (promoting creativity and innovation), idealised influenced (meaning charismatic role modelling) and individualized consideration referring to coaching and mentoring leadership style. 

In contrast to transformational leadership, transactional leadership refers to focus on employees ability to fulfil and achieve clearly defined goals (Hollander 1978; House 1971) and successful goal achievement is rewarded (Waldman et al. 1990). This exchange relationship between leader and employee is based on a contract of specified goals and emphasises on the process of achievement of objectives (Avolio and Bass 1988) but does not encourage employee's to develop their creativity and innovation skills (Jung 2001). Instead, employees are rather motivated extrinsically to perform their job under transactional leader but not expected to question and change the status quo in creative ways (Amabile 1998). 

Intrinsic motivation has been attached to creative and innovation performance of an individual as intrinsically motivated individuals usually prefer novel solutions, challenging status quo and trying out new ways for solving a problem at hand. (Amabile et al. 1994)

"Another common conception of creative leadership holds that the leader of creative people must inspire providing followers with a meaningful, motivating vision of the work and its implications (Sosik et al., 1999)."

Because leaders define the context in which their followers interact and work toward a common goal, we believe that previous findings of a positive link between transformational leadership and individual creativity can be extrapolated to an organizational level. 

Few studies have been made linking the transformational leadership and positive outcomes of employees' creativity in organisational level and outcomes, even though several studies have been made suggesting the positive relation between these factors. in their study Jung et al (2003) draw this link clearer and suggest that while leaders define the context and goals of their employes, transformational leadership can be extrapolated to an organisational level. 

Jung et al (2003) set four hypothesis. First of all, they suggest that transformational leadership is positively related to organizational innovation. Secondly Transformational leadership is positively related to employees' perceptions of (a) empowerment and (b) support for innovation. Thirdly
Employees' perceptions of (a) empowerment and (b) support for innovation have a positive relationship with organizational innovation.
Lastly, Employees' perceptions of (a) empowerment and (b) support for innovation moderate the relationship between transformational leadership and organizational innovation such that the relationship will be stronger when perceived empowerment/support for innovation is high rather than low.

"We have integrated extant discussions of leadership to propose four sets of hypotheses about how transformation leadership shown by top executives directly and indirectly affects innovation at the collective level of the organization" (Jung et al. 2003)
"Findings based on 32 Taiwanese companies provide support for our expectation that a direct and positive relationship exists between transformational leadership and organizational innovation. We also find positive and significant relationships between transformational leadership and empowerment as well as support for innovation and a positive relationship between support for innovation and organizational innovation. These results support our proposition that transformational leadership by the top manager can enhance organizational innovation directly and also indirectly by creating an organizational culture in which employees are encouraged to freely discuss and try out innovative ideas and approaches."

"Because undertaking innovative approaches to work typically requires making risky decisions, empowerment per se, if not accompanied by guidance and some measure of structure, could lead to negative consequences in a high power distance culture." (jung et al. 2003)

"Because many aspects of leadership behavior can be learned or modified, our findings suggest that organizations can improve their innovativeness by helping managers to develop and display transformational leadership behaviors through training and mentoring processes." (Jung et al. 2003)

"Second, and not totally independent of the possible role of power distance, our finding may underline a general need for transformational leaders to maintain a balance between letting people feel empowered and providing guidance via defining goals and agenda. Creative work processes often involve a complicated process of pulling available resources together to recognize current market trends, focus on the core message and strategy, and develop people in line with the strategy (Schein, 1992). Unless the leader plays an active role in providing guidance, coordinating and supporting these activities, employees or organizational units might wind up working at cross-purposes. Indeed, several studies have found a positive relationship between leaders' initiation of work structure and performance of creative activities (i.e., Keller, 1992). At the same time, the importance of allowing room for individual experimentation and initiative is emphasized by Mumford et al. (2002), who state that ??planning by leaders should not focus on the conduct of a specific piece of work. Rather, leaders' planning should focus on progress, the general types of projects that should be pursued, and the consequences of pursuing project results into development? (p. 716). " (Jung et al. 2003)

Time (Gruber and Davis 1988)
"Employees' willingness to experiment and take risks also may depend on the tightness of the resource and time constraints that they face at work. In turn, these aspects of the work environment may be affected by the extent to which superiors permit subordinate participation in establishing budgets and performance standards and in the latter's performance evaluation." (Jung et al. 2003)

encourage risk-taking (Sethi et al. 2001) 

"Sethi et al. (2001) were able to demonstrate that when a multifunctional team is given a powerful stimulus to take risks, the members of the team are more motivated to integrate their particular perspectives. As a result, members tend to identify themselves with the team (more than with their respective functional areas) and tend to be more committed to the success of the team. In this research, the cited interaction had a positive effect on the radicalness of innovation. "

Waldman and Bass 1991

Hohn, 2000

Amabile et al. 2004

Khaire, 2008


 Tolerance for ambiguity 
 
 Ambiguity is often perceived by individuals when lacking sufficient cues to structure a situation, and usually arises from novelty, complexity or unsolvability of situation at hand (Budner 1962). 
 
Role model
According to studies leaders have a strong direct impact on employee's behaviour and way of performing at workplace (Katz and Kahn 1978; Redmond et al. 1993)
Leadership plays a major role in defining group goals, controlling resources and providing rewards through interactive leadership process, making leadership behaviour an essential environmental variable in stimulating creative behaviour as a means for achieving goals. (tsekkaa tää rakenne vielä, Redmond et al. 1993) Katz and Kahn (1978) even refer to role of the leader in a sense where leader defines by his example the reality of workplace; norms, practices and culture. 

By defining organisational culture, climate and group norms leaders shape the way of working of employees. Through such role-modelling and mentoring process leaders also show employees in practise how tasks are performed. Employees, in turn, follow the example of leader in order to achieve high level of performance. (Redmond et al. 1993) 
 
Role-modeling stands also as powerful tool for opening employee's eyes and attitudes to new perspectives, thinking 'out of the box' and to adopt generative and exploratory thinking processes (Jung et al. 2003). 
"Role modeling by supervisors also can influence employee creativity" (Shalley and Gilson 2004) 
"Best thing a manager can do to encourage an inventing style is to act as role model and use that style herself" (Sternberg et al. 1997)
 
 Although different leadership styles and their effect on employee's creativity behaviour has not yet been studied widely, some studies show, how transformational leadership behaviour encourages employees look problems from different perspectives and thus widen their intellectual and creativity skills (Jung 2001; Sosik et al. 1998). Jung (2001) has studied the relation between leadership style and group creativity finding that transformational leadership is most likely to stimulate creative effort of employees. 
 
 In his study, Jung (2001) emphasises that transformational leadership skills can be practised in order to foster creativity and intellectual skills of employees and shape organisational culture. His study showed how transformational leadership; encouraging divergent thinking and solving problems at hand from unconventional perspectives, is likely to increase intrinsic motivation of employees leading to more creative problem solving and behaviour.  Through brainstorming activities that focus on non-traditional thinking and fantasising intellectual skills of employees can be enhanced (Sosik et al. 1998). 
 
Sosik et al (1998) furthermore suggested that anonymous ideating through nominal groups leads to better results and greater amount of ideas than brainstorming activities in real groups. This may be due to fear of failure and measurement of performance (Jung 2001) tsekkaa tähän vielä, tai yhdistä aiempaan. jossain se lukee.

"In this paper, we will break the silence surrounding these issues and expose the implications of Quin's message. We will introduce yet another paradox: innovation leaders should be controlled schizophrenics. We will argue that in order to managing innovation (processes) the responsible leaders (managers and/or consultants) have to behave and act in different and conflicting roles and take on different attitudes at the same time. They must do this without losing contact and rapport with their innovation team members. They need to have a great tolerance for dealing with the different conflicting and competing aspects of innovation within the innovation team. Like the leader, the innovation team is also struggling with the same conflicting aspects but with a different time scale, different interest levels and with differences in knowledge and experience. The innovation team members expect their leader(s) to be in charge and to be in control. Yet, they also want support, enthusiasm and trust. They do not want to hear that their leader sometimes has doubts about the direction to be taken, the sequence of the process and about the success of the project. If the innovation leader shows his doubts too strongly, then he ruins the game." (Buijs, 2007)

As Buijs (2007) argues, leaders who are to lead employees and work handling innovations need to understand the paradox and natural conflicts between routine processes (exploitation) in order to earn money in the present and the innovation processes (exploration) in order to earn money in the future. Buijs' (2007) four aspects for innovation (innovation process, psychological process of innovation team, creativity process, leading and playing), which leader should be able to master all providing a secure environment for a team to perform in novel and creative ways. As Buijs states "While the leader is already thinking about the next uncertain step to be taken, the team has to be encouraged to execute the present step comfortably."

"Incrementalism helps deal with the psychological, political, and motivational factors that are crucial to project success. By keeping goals broad at first, a manager avoids creating undue opposition to a new idea. A few concrete goals may be projected as a challenge. To maintain flexibility, intermediate steps are not developed in detail. Alternate routes can be tried and failures hidden. As early problems are solved, momentum, confidence, and identity build around the new approach. Soon a project develops enough adherents and objective data to withstand its critics' opposition.
As it comes more clearly into competition for resources, its advocates strive to solve problems and maintain its viability. Finally, enough concrete information exists for nontechnical managers to compare the programs fairly with more familiar options. The project now has the legitimacy and political clout to survive-which might never have happened if its totality has been disclosed or planned in detail at the beginning. Many sound technical projects have died because their managers did not deal with the politics of survival." (Quinn 1985)

"Of all of the
forces that impinge on people?s daily experience of the work environment in these organizations, one of
the most immediate and potent is likely to be the leadership of these teams?those ??local leaders?? who
direct and evaluate their work, facilitate or impede their access to resources and information, and in a
myriad of other ways touch their engagement with tasks and with other people." (Amabile et al 2004)

"The classic behavioral approach is the two-factor theory of leadership, which specifies that all leader
behaviors can be characterized as either task oriented (??initiating structure??) or relationship oriented
(??consideration??)(Fleishman, 1953). Task-oriented behaviors focus on getting the job done, and include
such things as clarifying roles and responsibilities, planning projects, monitoring the work, and
managing time and resources. Relationship-oriented behaviors focus on the socioemotional: showing
consideration for subordinates? feelings, acting friendly and personally supportive to them, and being
concerned for their welfare. It is important to note that in the leader behavior literature, the term
??support?? typically refers to relationship-oriented behaviors only, while in the creativity literature,
??support?? typically refers to all leader behaviors that could enhance creativity?both task- and
relationship-oriented. We adopt the latter, broader usage." (Amabile et al. 2004)

"This study suggests that a leader who interacts daily with subordinates may, through certain
behaviors directed at those subordinates, influence their daily perceptions, feelings, and performance,
ultimately influencing the overall creativity of the work that they do.--
In particular, local leaders display support for subordinates and their work by monitoring progress
efficiently and fairly, consulting with them on important decisions, supporting them emotionally, and
recognizing them for good work. They display a lack of support by monitoring progress inefficiently
or unfairly, giving unclear or inappropriate task assignments, and failing to address important
problems.--
The componential theory of organizational creativity (Amabile, 1988, 1997) provides a general
framework in which to understand these effects. It states that the perceived work environment can
have a significant impact on individual and team creativity, it identifies local leader support as one
important aspect of the perceived work environment for creativity, and it suggests some behavioral
elements that may constitute such support.--
The componential theory of organizational creativity (Amabile, 1988, 1997) provides a general
framework in which to understand these effects. It states that the perceived work environment can
have a significant impact on individual and team creativity, it identifies local leader support as one
important aspect of the perceived work environment for creativity, and it suggests some behavioral
elements that may constitute such support." (Amabile et al. 2004)

"Several behaviors deserve particular emphasis in the leader?s repertoire, behaviors requiring the
following: skill in communication and other aspects of interpersonal interaction; an ability to obtain
useful ongoing information about the progress of projects; an openness to and appreciation of
subordinates? ideas; empathy for subordinates? feelings (including their need for recognition); and
facility for using interpersonal networks to both give and receive information relevant to the project.
Perhaps just as importantly, there are also several behaviors for leaders to avoid or reduce, including
giving assignments without sufficient regard to the capability or other responsibilities of the subordinate
receiving them; micromanaging the details of high-level subordinates? work; and dealing inadequately
with difficult technical or interpersonal problems (whether due to technical incompetence, interpersonal
incompetence, inattention, or sloth)." (Amabile et al. 2004)

In their study, Shalley and Gilson (2004) present how leaders should use human resource practices in order to develop work context which improves the creativity skills of employees. 

"that many of the traditional roles of leadership are being redefined in today's flatter organizational structures. With employees having more direct responsibility over their day-to-day work, the leaders' role is being redefined so that they are more involved in external resource acquisition and boundary spanning. " (Shalley and Gilson 2004) 

"Oldham and Cummings (1996) found that supportive, noncontrolling supervisors created a work environment that fostered creativity. Finally, Tierney, Farmer, and Graen (1999) found that open interactions with supervisors and the receipt of encouragement and support lead to enhanced employee creativity."

 "However, the research on supervisor support consistently finds that contextual factors interact with individual characteristics to affect creative performance. Therefore, leaders need to understand their employees to provide the right levels of support needed for creativity to occur." (Shalley and Gilson 2004) 
 
 As Garvin et al. (2008) bring out in their article, reasonable question to ask for this fresh leadership approach is "where is the glory in being a "facilitator" as a manager? How do you get a management level made up of real humans who aspire to that role and will do it?" Elizabeth Long Lingo (2010) has offered one perspective to this question in her study with production of music. She claims that producer is the one bringing it all together; it is actually hard leadership exercise, where people from different fields and teams need to work together for one production, where there are no clear rules for who is controlling the output nor yardstick how good or bad the production is. Through creating a shared purpose and common goal in production team, and while still letting "other apply their distinctive expertise", a producer actually operates at the centre of the storm without being at the focus of attention as well as aims for productivity without being over controlling. According to this example, glory comes from being able to help others to find and realise their unique talents at the same time with achieving a collective goal. 
 
 "Team leader coaching is also likely to be an im- 
portant influence on team psychological safety. A team lead- 
er's behavior is particularly salient; team members are likely 
to attend to each other's actions and responses but to be 
particularly aware of the behavior of the leader (Tyler and 
Lind, 1992). If the leader is supportive, coaching-oriented, 
and has non-defensive responses to questions and chal- 
lenges, members are likely to conclude that the team consti- 
tutes a safe environment. In contrast, if team leaders act in 
authoritarian or punitive ways, team members may be reluc- 
tant to engage in the interpersonal risk involved in learning 
behaviors such as discussing errors, as was the case in the 
study of hospital teams mentioned above (Edmondson, 
1996). Furthermore, team leaders themselves can engage in 
learning behaviors, demonstrating the appropriateness of and 
lack of punishment for such risks." (Edmondson 1999)

"Organizations use a variety of types of teams. Team type 
varies across several dimensions, including cross-functional 
versus single-function, time-limited versus enduring, and 
manager-led versus self-led. These dimensions combine to 
form different types of teams, such as a time-limited new 
product development team or an ongoing self-directed pro- 
duction team. The team learning model should be applicable 
across multiple types of teams, because the social psycho- 
logical mechanism at the core of the model concerns people 
taking action in the presence of others, and the salience of 
interpersonal threat should hold across settings." (Edmondson 1999) en oo karma onko olennaista
 

\subsection{Practices and structures for developing and experimenting}

According to Amabile et al. (2002) clear time should be allocated for developing especially when the aim is to flourish idea generation, creativity, learning and experimentation of new concepts. Time pressure should be minimal in order bright ideas to glow as cognitive processing requires time of an individual and team. Yet, no sense of urgency leads employees easily to auto-pilot mode, in which routine tasks are performed without further thinking and analysing. Thus, creative time for playing with ideas, brainstorming, learning and experimenting should be allocated in an organisation in order truly new things to develop. Shared goals are once more essential in engaging team members to play with ideas and feel more motivated in developing their work.

Creativity, exchanging ideas and turning them into action requires intrinsic motivation from employees (Jung 2001). Thus, in order to increase creativity and innovation at workplace, leaders should foster organisational culture in which individuals find their motivation in divergent thinking and trying out new ways of performing tasks (Amabile 1998). 

Mumford et al (2002) -artsusta
-  In the sense that leaders must acquire resources and encourage the generation of new ideas (e.g., McGourty et al., 1996, tää lisätty referensseihin), there is some truth to this view. However, leaders seem to serve a number of other roles when people are engaged in creative work: evaluating their ideas (Sharma, 1999), integrating their ideas with the needs of the organization Mumford, 2000a and Mumford, 2000b, and creating conditions where people can generate ideas in the first place (Andrews and Gordon, 1970, tämä löytyy myös referensseistä). 


 "Amabile (1998) also has suggested that by influencing the nature of the work environment and organizational culture, leaders can affect organizational members' work attitudes and motivation in their interactions, thereby affecting their collective organizational achievement."
 
 "The extent to which they
will produce creative?novel and useful?ideas during their everyday work depends not only on their
individual characteristics, but also on the work environment that they perceive around them" (Amabile et al 1996)

Prior studies show how creative efforts of employees require sufficient amount of time and energy. (Gardner 1988; Getzels and Csikszentmihalyi, 1975) Redmond et al. (1993) argue how leaders should allow enough time for problem solving and creative actions. 

"For instance, while job rotation has become popular, managers need to ensure that employees have enough experience in an area of work if they want them to be creative. Therefore, while individuals from different areas may bring a new perspective to the work, they also need to have sufficient experience and familiarity with the target area so that creativity can occur." (Shalley and Gilson 2004) 

"We recognize that managers are sometimes wary of giving employees too much autonomy, such as giving them full control over how their work is planned and conducted. However, giving appropriate levels of autonomy to employees may be useful." (Shalley and Gilson 2004) 

"Taken together, this body of research indicates that leaders need to set appropriate goals and requirements so that individuals will aspire to be creative. At the same time, they need to be aware that if job requirements or goals are set for behaviors that may not result in creative outcomes or may directly contradict engaging in creative activities, then employees may exhibit less creativity in their jobs." (Shalley and Gilson 2004) 

"when managing for creativity, time is a critical resource that managers need to ensure their employees have access to. Here, it is important to emphasize that, in general, it is far easier for most employees to stick to routine tried and tested methods that are typically more efficient rather than experimenting and trying to come up with creative approaches. For instance, it takes more cognitive effort to generate multiple alternatives, suspend judgment, and look at problems in a different and often divergent manner. However, by engaging in creative activities, the quality of decisions or judgments should be better." (Shalley and Gilson 2004) 

"Amabile and Gryskiewicz (1987) found that one frequently mentioned factor necessary for promoting creativity was sufficient time to think creatively, explore different perspectives, and play with ideas.  Likewise, Katz and Allen (1988) found that for engineers working on new technologies, uninterrupted time was considered to be critical."

"A recent study by Amabile et al. (2002) found that individuals under time pressure are significantly less likely to engage in creative cognitive processing."

"Therefore, managers need to ensure that employees have enough time to be creative, which can be especially difficult in today's fast-paced, rapidly changing world." (Shalley and Gilson 2004) 

"In addition to time, employees need access to material resources to be creative (Katz and Allen, 1985). However, with regards to material resources, managers are faced with an interesting dilemma. That is, while material resources have been described as important for creativity, it also has been suggested that their availability or abundance might negatively impact creativity (Csikszentmihalyi, 1997). For example, while resources are needed to perform one's job, not having everything that is needed readily at hand, in fact, may stretch employees to think of different ways of doing their work. In other words, a lack of material resources may actually help foster creativity. Taking this a step further, Csikszentmihalyi (1997) suggests that resources can make individuals too comfortable, which can have a ?deadening effect on creativity? (p. 321). "

"Providing employees with performance feedback is a key function that many managers struggle with. Giving feedback can be particularly important for creativity and yet particularly difficult in that creativity often involves trying new things and taking risks. " (Shalley and Gilson 2004) 

"Another important component of organizational structure is how levels of responsibility and formal reporting relationships are organized. For instance, a highly bureaucratic organization may not encourage employees to try new ways of doing their work, whereas a flatter structure with wider spans of control may be more conducive to employee creativity." (Shalley and Gilson 2004) 

"Cummings and Oldham (1997) found that individuals with creative personalities produced more creative outputs than those with less creative personalities only when they were surrounded by an organizational context that facilitated creativity."

"With regards to creativity, it should be important that employees perceive their work context as one where decisions are made and applied in a just manner. " (Shalley and Gilson 2004) -> Fariness

"While overall climate is often regarded as a hard thing to change, there are several components of climate that are reasonably manageable and should have an effect on creativity. For example, fostering a climate where risk taking and constructive task conflict are encouraged can be role modeled and actively encouraged and supported by management. Likewise, a review of a division's or organization's management hierarchy and reporting structure may highlight that employees are not encouraged to make decisions on their own and thus may be less likely to try new ways of doing their work. Finally, if the bureaucracy associated with changing anything is such that it takes a great deal of time and effort to get new ideas considered, employees also may be less likely to try new approaches to work." (Shalley and Garvin 2004) 

"The human resource practices used to select, train, appraise, and reward employees all need to be systematically linked together so employees know what is expected of them and when and how. This also ties back to the importance of procedural justice in that if employees understand how, when, and for what they will be rewarded, promoted, or even fired, then they should have a stronger sense of fairness and subsequent organizational commitment, loyalty, and increased levels of organizational citizenship behavior. In addition, it is specifically these types of attitudes that need to be fostered for creativity to occur. For instance, employees who are not loyal or committed to their organizations will not be willing to give more than is required by their job and therefore will be more likely to stick to the tried and true ways of performing their tasks rather than searching for alternative solutions." (Shalley and Garvin 2004) 

"The practical implications of our review for the day-to-day management of creative people should be highlighted. First, across the empirical studies reviewed, one common theme is that individuals need to feel they are working in a supportive work context. This applies to how leaders interact with employees, how coworkers, team members, and even others outside of work interact with employees, whether sufficient resources are available, how employees expect to be evaluated and rewarded, and whether the climate is perceived to be supportive (e.g., a perceived fair environment). Thus, managers should attempt to increase the supportiveness of the work context." (Shalley and Garvin 2004) 

Amabilen (1997) tekstissä pdf:n sivulla 17 on hyvä kiteytys Amabilen osalta, mitä innovointi vaatii organisaatiolta. 

In addition, Amabile and Khaire (2008) argue how sufficient time and resources should be allowed for exploration. 
 
\subsection{Attitude towards failure and risk}

"those in a 
position to initiate learning behavior may believe they are 
placing themselves at risk; for example, by admitting an er- 
ror or asking for help, an individual may appear incompetent 
and thus suffer a blow to his or her image. In addition, such 
individuals may incur more tangible costs if their actions cre- 
ate unfavorable impressions on people who influence deci- 
sions about promotions, raises, or project assignments" (Edmondson 1999)

"Asking for help, admitting errors, and 
seeking feedback exemplify the kinds of behaviors that pose 
a threat to face (Brown, 1990) even when 
doing so would provide benefits for the team or organization (edmondson 1999)"

"there is no innovation process without failures and mistakes. Organizations need to learn from them as quickly as possible. If the organization out-learns its competitors, they then take the lead. And taking the lead is what innovation is all about!" (Buijs 2007)

"The phenomenon of threat 
rigidity has been explored at multiple levels of analysis, 
showing that threat has the effect of reducing cognitive and 
behavioral flexibility and responsiveness, despite the implicit 
need for these to address the source of threat (Staw, Sand- 
elands, and Dutton, 1981)."

"In sum, people tend to act in 
ways that inhibit learning when they face the potential for 
threat or embarrassment (Argyris, 1982)." 

"Nonetheless, in some environments, people perceive the 
career and interpersonal threat as sufficiently low that they 
do ask for help, admit errors, and discuss problems. Some 
insight into this may be found in research showing that fa- 
miliarity among group members can reduce the tendency to 
conform and suppress unusual information (Sanna and Shot- 
land, 1990); however, this does not directly address the 
question of when group members will be comfortable with 
interpersonally threatening actions." (Edmondson 1999)

Failing is widely considered as essential part of learning (Farson and Keyes, 2002; )

Andriopoulos and Lowe (2000) :
"The risk attached to creative work implies both a need to experiment and a need to tolerate failure Andriopoulos and Lowe, 2000 and Quinn, 1989."(Tää andriopoulos löytyy PDF:nä, käy läpi, on varmaan hyvää settiä!) 

Suoraa lainausta Mumford et alilta (2002) 
"Thus, creative work is contextualized with the success of creative ventures depending on an awareness of the capabilities of, and pressures on, extant socio-technical systems. In fact, it is this contextualization of creative activities that accounts for such well known phenomena as simultaneous invention and the tendency for innovation to occur in spurts within a given industry Csikszentmihalyi, 1999"

Feeling of self-efficacy may affect individual's willingness to provide unique and novel ideas even when some degree of risk is involved (Mumford and Gustafson 1988). Training, coaching, giving feedback and assigning tasks seem to be useful approaches for leaders, who pursue to contribute empoloyee's self-efficacy (Amabile 1983; 1988). 

Fear of failure can be decreased through transformational leaders who foster the culture of intrinsic motivation and rewards from creative endeavours, sea exchange and discussion. (Amabile 1998) 

According to Amabile et al.(1996) creative solutions in an organisation can be achieved by encouraging employees to reach and experiment new perspectives and ways of performing. Essential for this is not being punished for negative outcomes. Organizational environment that allows failing is likely to assist in employees acquiring diverse perspectives and questioning the status quo and habitual way of performing. 

"In addition, established theory recognises the need for developing opportunities where employees can exploit uncertainty. Sternberg et al. (1997) argue that managers should let messiness exist. In other words, they suggest that uncertainty associated with creative projects must not be controlled in order to establish some order."
"The importance of
challenging work has also been emphasised
by Amabile (1997), who states that matching
creative employees to assignments, based on
their skills and interests, enhances their
motivation toward work." 

"Motivate people to contribute ideas by 
making it safe to fail. Stress that the goal is to 
experiment constantly, fail early and often?
and learn as much as possible in the process. 
Convince people that they won?t be punished 
or humiliated if they speak up or make 
mistakes." (Amabile and Khaire 2008)

"The comparison of perpetual
challenging with established corporate
creativity theory reveals that the emergent
category of adventuring comes closer to the
five stages of Amabile's (1988) componential
framework of creativity. Specifically, the
properties of adventuring, such as
introspecting, scenario making and
experimenting can be associated with
Amabile's preparation, response generation
and validation stages of the creative process.
Moreover, her conceptual model emphasises
the fact that the process can have a negative
consequence, which can also be found also in
the emergent category of ``adventuring''
under the title ``mistake making'' within this
grounded theory study." (Andriopoulos and Lowe, 2000)


"to develop new and useful products or processes, individuals have to be willing to try and to possibly fail. For many, this is not an easy thing to do and can, in part, depend on the individual's predisposition toward risk as well as the organizations culture, which will be discussed later in this article. " (Shalley and Gilson 2004) 

"Research has indicated that people tend to avoid risk and prefer more certain outcomes (Bazerman, 1994). However, because creativity does not just happen but rather evolves through a trial-and-error process that involves risk taking, failure will often occur along with success. If employees are risk averse, it is much easier for them to continue performing in more routine ways rather than take a chance with a new, and potentially better, approach. Therefore, a key in the motivation of employees toward creativity is to ensure that they feel encouraged to take risks and break out of routine, safe ways of doing things." (Shalley and Gilson 2004) 

"If leaders value and want employees to be creative, a critical contextual factor they need to attend to is fostering an environment where risk taking is encouraged and uncertainty is not avoided. This has been referred to as providing a culture where employees feel psychologically safe such that blame or punishment will not be assigned for new ideas or breaking with the status quo (e.g Edmondson, 1999). In support of these arguments, Nystrom (1990) found that organizational divisions were more innovative when their cultures reflected challenge and risk taking, " (Shalley and Gilson 2004) 

"participative safety, being able to give input without being judged or ridiculed, has been positively linked to creativity (De Dreu and West, 2001)"

Predicting the future being impossible, focus should be in managing risks involved in playing with creative ideas in both the company and individual level. As Stenberg et al. (1997) state, "as uncomfortable as it is, while not being able to predict and control uncertainty in creative projects, the messiness does have to let exist. Kanter (1983) continues that, actually, opportunities grow from uncertainty and creative endeavours rise when struggling with uncertainty and mess, as individuals impose order where it does not exist, and thus individuals are forced to form new connections. Furthermore, allowing employees freedom to act actually arouses desire to act. (Kanter 1983)

According to Amabile and Khaire (2008) essential part of creating a safe environment for creativity is managers to decrease the fear of failure. Instead, constant experimenting should be the goal of working, learning by doing and iterating until sufficiently is learnt from the process. 

Furthermore, when company grows, it usually leads to more conservative actions and increase in fear of failure. When fearing failure managers tend to deny it happened and erase it from the memory instead of learning from it. (Amabile and Khaire 2008) Tämä näkyi jossain haastiksessa! 

According to Edmonsdon  (1999) "any business that experiments vigorously experience failure?which, when it happens, should be mined to improve creative problem solving, team learning, and organizational performance" 

"the premise that learning behavior in so- 
cial settings is risky but can be mitigated by a team's toler- 
ance of imperfection and error. This appeared to be a 
tolerance (or lack of tolerance) that was understood by all 
team members--
The implication of this result is that people's beliefs about how oth- 
ers will respond if they engage in behavior for which the out- 
come is uncertain affects their willingness to take 
interpersonal risks.. " (Edmondson 1999)

By creating an environment that serves psychological safety for employees, organisations may capitalise on failure. In a safe environment employees are convinced they are not humiliated or punished when failing or saying out loud their ideas, concerns or raise discussion. Furthermore, failure can be divided in three separate categories: unsuccessful trials, system break-downs and process deviations, which all need to be recognized and analysed and dealt with in order learning to happen. Especially unsuccessful trials are fruitful and essential for creative learning, yet overcoming "deep ingrained norms that stigmatise failure and thereby inhibit experimenting" is needed. (Garvin et al. 2008)
\end{document}  