\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\setcounter{secnumdepth}{5}

\begin{document}

The emphasis of the study was on recognising factors affecting experimentation behaviour. The results suggest that various factors have an effect on experimentation behaviour and combined to literature review experimentation-driven approach can be considered as one form and style of learning and developing an organisation. 

\section{Experimentation-driven innovation}
Failed experiments should not be considered as failing, instead they offer valuable learning points. 

According to Thomke (2003), in the beginning every product is an idea, that was being shaped through the process of experimentation, and the ability to do experimentations is actually a measurement of company?s ability to innovate.Experimentation is essential in order to learn about the idea, concept and prototype and whether it actually addresses a new need or a problem or solves the one at hand. Prototyping is critical part of the process, as testing the prototype in a real environment gives instant and valuable feedback for further development. Thomke (2003) suggests four steps for organizations to be more innovative. First of all, organization should allow and manage the work for the employees so that fast experimentation is possible. This usually requires challenging routine ways of working and shaping the routines, yet fast experimenting is essential in order to get rapid feedback for shaping the ideas. In addition, team engagement is essential, as the whole team need to understand the meaning of experimenting and developing and it should be encouraged to sharing information and ideas in as early stage of development process as possible and throughout the process. Thomke suggests using small teams and parallel experiments especially when the time is the most critical factor. 

Secondly, failing early and often, yet avoiding mistakes is important for experimenting. Failure can disclose important information and reveal gaps in knowledge, and is thus important as early phase of the development as possible. However, according to Thomke (2003), this is not an usual way for an organisation to think about failure, thus building the capacity for rapid experimentation as well as tolerating and learning from failure is essential and often requires overcoming ingrained attitudes. Encouraging and creating a culture where failing is allowed and not being afraid of, brainstorming sessions where judgement is not allowed are important for 

However, Thomke (2003) does not suggest failing and making mistakes as a result of poorly planned experiments. Mistakes and failures produce most value, when the experiment is well planned and the goal or hypothesis that needs to be tested is clear.

Thirdly, anticipating and exploiting early information can save a lot of resources in the development process. If problems are shown in the late-stage of the process, they can be even 100 times more costly than the ones discovered in the early stage. According to IDEO, an innovation and design-firm, using human-centered design-based approach, the key elements in the design process and prototyping is it being rough, rapid and right. The right-element reminds that even though the prototype itself is likely to be incomplete, it has to show the right specific aspects of a product. This forces developers to decide the factors that can initially be rough and those that must be right. In addition, exploiting early information serves as a good method for developers reflecting changing customer preferences. Briefly, information in the early stage of the developing process should be listened and discovered carefully, as the problems are cheaper and easier to solve. 

Lastly, for enlightened experimentation Thomke (2003) puts emphasis on combining new and traditional technologies. 

Check next:
Gerber and Carroll, 2011: The psychological experience of prototyping
Gerber 2009: Prototyping: facing uncertainty through small wins
Bandura, 1997. Deci and Ryan, 1985: Intrinsic motivation and self-determination in human behavior. 
Schunk, 1984: The self-efficacy perspective on achievement behavior
Floyd, 1984: A systematic look at prototyping
Gilovich, 1991: How we know what isn?t so: The fallibility of human reason in everyday life
Weick, 1984: Redefining the scale of social problems
Cross, 2002: Creative cognition in design: processes of exceptional designers
Schunk, 1984: The self-efficacy perspective on achievement behavior?
Taylor and Brown, 1988: Illusion and well-being: a social-psychological perspective on mental health
Deci and Ryan, 1987: The support of autonomy and the control of behavior. 
Seligman, 1990: Learned optimism
March, 1991: Exploration and exploitation in organizational learning
Hackman and Oldman, 1975: ?Motivation through the design of work?, 1980: ?Work redesign

\section{Change and learning in an organisation}

\subsection{Learning as competitive advantage and way to survive}

Understanding change analytically in the turbulent world appears challenging. Change being hectic and fast calls for different skills and strategy than before. Only when change is understood can it be managed, and in order to survive new perspective and understanding towards change is required from an organisation. 

In the changing environment tolerance for uncertainty is needed, and while future can not be predicted, forecasting is a usable method in order to cope with the anxiousness resulting from the uncertainty. Furthermore, even more emphasis should be put on ability to learn and adapt to changes. \subsection{Change in organisational level}
Hammer and Champy (1993) have summarised aspects of change in organisational environment, beginning from the change in organisational structure; from functional departments to process teams. Work tasks change from simple and detailed tasks to multi-dimensional knowledge work while employees are becoming more autonomous instead of strict control. Furthermore, instead of educating, focus is in the learning of an employee, and evaluation of work will change from operations to outcomes. Knowledge and capability are preferred over single performance and values change to more productive behaviour than over-protective. Superiors turn from leaders of the work to coaches and hierarchical organisational structures turn lower while managers focus on leadership instead of task management. 

According to Arie de Geus (1997) the only one can maintain company's competitive advantage is to make sure the company is able to learn faster than rivals. Generally organisations are considered as machines, yet recently more emphasis has been put on organisations as living organisms. When considered as machine, organisational model is mechanic and simple, which purpose is to gain profit. Whereas, organisation as a living organism is a whole-systemic model, and organisations are considered as place which has deeper, permanent meaning offering people the opportunity to grow and fulfil themselves while earning money.

Liable vision of the future focuses on the latter perspective of organisations, where learning and renewal form the essence of being. 

According to Huy and Minztberg (2003), organisations learn best through small experiments and trying out new things, and the closer and more related experimentations are to customers and customer interfaces, the more can be learned. 
\section{Organizational learning}

From the data rose various factors affecting experimentation behaviour, and the literature review revealed how similar factors are related to learning in an organisation and organisational learning. 

Through the understanding of organisational learning real growth and support for learning can be offered. Learning process needs to be understood all from organisational, team and individual perspective. 

Organizational learning is approached from two different perspectives in literature. On the one hand, learning is considered as an outcome, and on the other it is considered as a process (Edmondson 1999). In the first perspective organizational learning is referred to be an outcome of a process of organisations "encoding interferences from history into routines that guide behaviour (Levitt and March 1988). Whereas process perspective define learning as a process of continuous trial and error (Argyris and Schön 1978). Further definitions of learning are presented in chapter zz. In this thesis, learning is considered as the latter tradition of learning, which allows the growth and improved performance of individuals and organisations.

Organizational learning differs from individual or team learning. Organizational learning occurs through the shared knowledge, insights and approaches of the employees of an organisation. Secondly, organisational learning is based on prior knowledge and experience, the memory of organisation, which consists of the ways of working, processes and instructions of an organisation. Even though individual and team learning are highly related to organisational learning, it is not the sum of the previously mentioned. (Sydänmaanlakka 2007)

Various definitions of learning organisations have been presented. One of the most famous definitions is from Peter Senge (1991), who describes learning organisation as follows: "Learning organisation is an organisation, where people are able to constantly develop and achieve intended results; where new ways of thinking are born and where people share goals and learn together."
\subsection{What is learning?}
Generally, learning process consists of experience, evaluation, understanding and applying new knowledge into use. Sydänmaalakka (2007) claims clarifying and understanding phases of learning process assists in supporting learning and organisational recurrence. 

Learning is changing, growing, evolving and developing. Generally time for reflection is needed in learning process, when new information or experience is reflected and compared to prior knowledge and experiences and attached to previous frameworks, and when considered useful, new experiences and knowledge is combined to previous knowledge providing new way for performing in the future. (Sydänmaanlakka, 2007)

Various definitions of learning have been suggested. Sydänmaanlakka (2007) defines learning as a process, where an individual acquires new information, abilities, attitudes, experiences and contacts, which result changes in his actions. According to this definition, learning is a process which is affected by individual's cognitive, emotional and physical skills. Information acquirement always concerns of individual's interpretation, thus learning is not mechanical action but making meanings through interpretations. Furthermore, learning is not only acquiring new information; the process involves abilities, attitudes, emotions and values. In addition, all experiences are essential in learning, and the more experiences individual has, the easier it is for him to learn from the new experiences. In order to truly understand, several skills even require prior experience from the field. Furthermore, contacts and networks serve as significant factors in learning process. True learning consists of adapting new information and processes resulting in change in action, behaviour or thinking. According to Kolb (1984) learning is a process that creates knowledge through transaction of subjective and objective knowledge of an individual. Furthermore, 

In addition, motivation is essential part of learning (Sydänmaanlakka, 2007)

\subsection{Levels/stages of learning}

Learning occurs in several levels. In the bottom level is the belief of knowing something. Already a step further is the comprehension of not knowing something, right before the level of knowing, which consists of acquiring new information, following the step of understanding. Understanding consists of actual comprehension of phenomena at hand, and this level also concerns the attitudes and emotions of an individual. More time is required for actually understanding something than only acquiring new information. Level of applying proves how recently learnt things is successfully applied in practise. The level of applying requires management of previous levels - knowing and understanding. According to Sydänmaanlakka (2007) true and deep learning comprises all these levels and enables developing and renewal of already known things and approaches. 

Learning types
Briefly, learning styles of individuals, teams or organisations can be divided into four categories. 
Reactive learning
Proactive learning
Action learning
Questioning learning

Kolb's (1984) general model for learning process begins in experiences of an individual and the ability and motivation to learn from them. Motivation stands for key to learning, following time for reflecting and processing new perspectives to previous knowledge and experiences. After reflection the knowledge is processed and understood, following the final stage of applying knowledge in practice. Learning can best be supported when all these phases are taken into account. 

Kolb's learning model proposes a four-stage cycle where learning is seen as continuous, interactive process. Those stages are concrete experience (CE; experiencing), abstract conceptualisation (AC; thinking), active experimentation (AE; doing), which involves active trial-and-error learning, and reflective observation (RO; reflecting) in a different manner depending on an individual.  Honey and Mumford build on this model in their learning style questionnaire and divide learners according to Kolb to activists, reflectors, theorists and pragmatists. (Cassidy)

Factors enhancing learning process consists of motivation, joy of success, experimenting and documenting. 

Based on the learning process of Kolb, several models have been developed to describe different learning styles. One is Honey and Mumford's model, where individuals are divided into four categories based on their learning style: activist, reflector, theorist, and pragmatist.

\subsection{Learning and change from team perspective}

In order to create new value and competitive advantage in rapidly changing and uncertain organisational environments, new managerial imperative is growing focusing on teams. Thus, supporting teams in their work and understanding the aspects of team learning is required. (Edmondson 1999) Recent studies has moved the focus from individual learning to team learning

Fast-pace environment requires organisations to enhance the ability of teams to learn and create environment where learning is safe and possible. Uncertainty affects all working life, change is faster and job security lower. Psychological safety for individuals at work can be increased through great teams and teamwork. Employees should be engouraged to ask questions, seek for help and tolerate mistakes and uncertainty. (Edmondson 1999)

As Edmondson pointed out, psychologicl safety as a means to promote team performance is increasingly relevant both in future work and research. 

Definition of team consists of a small group of people who have supplementary knowledge and abilities compared to each other, and who share a goal, targets and way of working and approach. Team also shares a feeling of being responsible for its performance. (Edmondson, 1999)

Team opportunities and full advantage of team effort can be threatened by demanding schedules, long-standing habits and unwarranted assumptions. In addition, in order to function, team needs a clear purpose and vision what makes it a team and why it exists. Teams get energy from significant performance challenges regardless of where they are in the organisation. Set of shared, demanding performance goals usually form a team, and personal chemistry or willingness to form a team may boost that, yet teamwork is essential for an actual team. Indeed, in order to receive great results teams should focus on performance regardless of the organizational hierarchy or what team does. Thus, team performance may exceed the results of what could be achieved if employees were acting alone as individuals without the team effort.  (Katzenbach)

Organizational leaders play a great role in establishing strong team performance culture. This can be achieved through addressing and demanding performance that meets the need of the customers, employees and shareholders. Teams should not be fostered by the sake of the team only, rather should leaders clearly state how the team performance affects to customers and through that clearer performance ethics and cultures are fostered. In addition, even though people tend to have great sense of individualism, it does not have to bias the teamwork performance, as real teams find ways to support individual strengths and performance for shared goal. Furthermore, in order to team function properly and efficiently, discipline across the team and organisation is needed, focusing again on performance. "Groups become teams through disciplined action." According to Katzenbach great team performance consists of continuos work of shaping a common purpose, agreeing on performance goals, defining a common working approach, developing high level complementary skills and being transparent on the results.  (Katzenbach) 

Team psychological safety refers to Amy Edmondson's concept of team members shared belief team being safe for interpersonal risk-taking. Together with team efficacy these have great affect on team performance and learning in an organisational team work. According to Edmondson, team performance and outcomes can be shaped through team structure and shared beliefs. For instance, employee's willingness to take interpersonal risks depends highly on the experience of team safety and person's beliefs how others will respond in ideas or situations involving uncertainty. Team psychological safety refers to interpersonal trust among team but beyond that mutual respect and caring.  (Endomndson 1999)

According to Edmondson (1999) learning behaviour consists of seeking feedback, sharing information, asking for help, talking about errors and experimenting. These factors were also recognised from the data as factors affecting experimentation. Thus, experimentation behaviour seems to relate to learning of an individual, teams and organisations and can be supported by supporting these factors. 

Studies have shown the essence of psychological safety in teams, defining how structural and interpersonal characteristics both have effects on learning and performance in teams. Psychological safety can be boosted for instance through structural factors such as context support and team leader coaching affecting behavioural and performance outcomes. Furthermore, climate of safety and supportivenesss encourages employees to seek for feedback and ask for help in addition to admit and reflect mistakes.  (Edmondson 1999) 

As Edmondson suggested, studies in real work teams are required, and this study aims to add on this aspect studying factoring affecting experimentation behaviour in real work context. 


\subsection{Individual learning}

According to experiential learning theory, learning is "the process whereby knowledge is created through the transformation of experiences." (Kolb) Experiential learning process stands as a way describe the central process of human adaptation to the social and physical environment. (Kolb) Learning involves total concept of human being - feeling, thinking, perceiving and behaving. (Jung, 1923) Learning should be conceived as a holistic adaptation process providing bridges across life situations and underlaying the lifelong process of learning. (Kolb)

However, learning may not be the first thought in mind when considering an immediate reaction to the situation or problem at hand, rather is it considered as performance. Furthermore, long-term adaptations to our previous experiences and beliefs is mainly considered as developing, not learning. Thus, when talking about developing an individual or an organisation, the question highly concerns and is related to learning. (Kolb)

According to Kolb (..) experience plays a central role in learning, and learning is a process consisting of experience, perception, cognition and behaviour. Immediate experience forms a basis for reflection and observation, following assimilation to a theory from which new implications for action are deducted. In order to create new experiences, these implications serve as guides. 

This Lewinian approach emphasises here and now -concrete experiments in testing and validating abstract concepts. Thus, experience of an individual is a focal point of learning, and giving personal meaning to abstract concepts, which can be afterwards shared with others. Furthermore, receiving feedback is considered essential in this approach for learning, as it serves a continuous process for goal-oriented action following evaluation of that action. Feedback can thus boost effective, goal-oriented learning process. (Kolb)

Kolb presents Dewey's model of Learning, which adds to Lewinian model how through learning the impulses, feelings, and desires of concrete experience into higher-order purposeful action is transformed. Both Lewin and Dewey consider learning as a dialectic process integrating experiences and concepts, observations and actions. 

Piaget has formed in 1970 model of learning and cognitive development, which presents similar aspects to Lewin, Dewey and Kolb. Piaget also stated how learning is an interactive process between individual and environment. Learning is a mutual process between accommodation of concepts or schemas to experiences around us and assimilation of events and experiences into existing concepts and schemas. This intelligent adaptation, learning, results from the tension between accommodation and assimilation. Through this tension growth and higher-level cognitive functioning occurs. 

Kolb summarises above models. First of all, learning should not be conceived in terms of outcomes, but it is best conceived as a process. Ideas are not fixed and and immutable elements of thoughts, but can be formed and re-formed through experience,  

Bringing the experiential learning into education implications, all learning can be considered as relearning. Thus, all learning situations should take into account people arriving from all different experiential backgrounds to what they build their new experiences and knowledge on. This partly explains very likely resistance to new ideas, as when new information and experiences are in contradiction to old beliefs and experiences, new ideas and information is more difficult to adapt. In the education process learner's old beliefs and theories should be brought out, examined and tested, following integration of the new models and refined ideas into learner's belief systems. (Kolb)

According to Piaget individual learning and adaptation of new ideas occurs through integration or substitution, integration leading to stronger part of learner's conception of the world, whereas substitution requires real questioning of previous conceptions, and thus might take longer for the learner to adopt. 

According to Kolb, learning is a process filled with tension and conflict, and new knowledge, skills and attitudes are achieved through experiential learning, which consists of four modes and required abilities of learners: concrete experience abilities, reflective observation abilities, abstract conceptualisation abilities and active experimentation. First of all, individuals must openly involve themselves in new experiences, reflect and observe them from various perspectives, create concepts that can be integrated into more abstract theories as well as they need to be able to use these reflections and theories in active daily decision-making and problem-solving. Quite a task and ability list for individuals, and this is to show how learning requires various abilities from different perspectives and to choose which learning perspective to use in a situation at hand. Furthermore, Kolb divides the learning into two dimensions: "first of all representing the concrete experiences at one end and abstract conceptualisation at the other. " The other dimension includes active experimentation at one end and reflective observation at other. Thus, the level of learning depends on the way in which conflicts ... Kesken, kolb page 15. One mode dominating the conflict resolving process leads to learning dominated by that mode. However, the aim is to create a synthesis between the different modes in order to boost the process of learning and personal development. 

Also Sydänmaanlakka emphasises different learning styles and modes, which should be understood and used variously in learning and problem-solving situations. 

In addition the meaning of environment in learning should be emphasised. Learning concerns of transaction between an individual and the environment, learning does not happen only inside of individual's thoughts, experiences and processes but is dependent on the real world environment. 


\subsection{Factors hindering experimentation and learning in an organisation}

Several factors may affect on the gap between idea and action in employees of an organisation. 

One essential factor is the beliefs, emotions and actions of an employee. For instance, when feeling under risk or pressure, an employee is likely to inhibit learning behaviour as a result of feeling the fear of being rejected or feeling they are placing themselves at risk. This may occur in a situation where an employee should ask for help, yet is afraid of admitting he lacks abilities, skills or knowledge. (Endomndson 1999). In addition, admitting mistakes, asking for help and seeking feedback are all relevant abilities in the recent organisational world, yet threatening for an individual's image of himself and his skills (Brown 1990). Indeed,  learning is often inhibited as people tend to act in a way that inhibits learning when facing the potential for embarrassment of threat, even though their transparency and honesty would be highly important for the behaviour of the team. (Argyris 1982)

The phenomenon of threat of employees in organisations is widely studied and consensus is rising how threat effects on cognitive and behavioural flexibility and responsibility in reducing manner. 
\subsection{Building blocks: How can organisational learning be supported?}

According to Garvin et al. (2008) in learning organisation employees excel at creating, acquiring and transferring knowledge. They define building blocks for learning organisation: supportive learning environment, concrete learning processes and practices and leadership behaviour that reinforces learning. Building blocks can be considered and measured as independent components yet each of them vital to the whole. In order to improve long-term learning of an organisation, strengths and weaknesses of an organisation and its unit needs to be recognised. 

In order experimenting to happen in an organisation safe and supportive environment has to be created. According to Garvin et al. (2008) supportive learning environment consists of four characteristics: psychological safety, appreciation of differences, openness to new ideas and time for reflection. 

Psychological safety means that learning of employees occurs when employees do not fear being rejected, ask naive questions, make mistakes or present viewpoint of minority. Psychologically safe environment enables employees comfortably to express their thoughts at work. Appreciation of differences is important as opening minds for different ideas and world views increases both energy and motivation, brings out fresh thinking. Learning occurs when employees become aware of opposing ideas in a safe environment, and additionally openness to new ideas is required. Novel approaches are relevant for learning, thus employees should be encouraged in risk-taking and exploring and testing uncertain things. Lastly, through providing time for reflection learning in safe environment occurs. Instead of looking and judging by numbers of hours of work or results employees should be given enough time to reflect their work. Analytic and creative thinking will not occur under stress, heavy workload and too tight schedule. Under stress ability to recognise and react to problems and learn from experiences deteriorates. In supportive learning environment time for reflection is allowed. 

Similar characteristics were found in data and are further presented in chapter xx. 

Edmondson in her research (1999) studied whether beliefs in interpersonal context varied between the same organisation and whether it affected in team outcomes

According to Garvin et al. (2008) second building block of organisational learning, consists of concrete learning processes and practices. It includes experimentation, information collection, analysis, education and training and information transfer. Organizational learning can be supported trough concrete steps and activities which are tested and further developed through experimentations. Furthermore, information and intelligence about customers as well as technological trends should be collected systematically and further analysed focusing on identifying problems and solving them. Training and education of new and established employees is an essential part of practices and processes. Finally, through transparent and meaningful knowledge sharing organisational learning can be enhanced, focus being on clear, well-defined and working communication systems that employees can easily relate and feel useful. Concrete processes together with efficient knowledge sharing methods ensures that essential information is available quickly and efficiently for employees who need it. 

 Thus, emphasis should be put on creating and defining concrete learning processes and practices.
 
Thirdly, leadership behaviour should reinforce learning. Behavior of leaders is highly related to the performance of employees (Kim and Mauborgne 2014) and organisational learning (Garvin et al. 2008). In order to encourage employees to learn, leaders should prompt dialogue and debate, ask questions and listen to employees. Example of the leaders was also recocgized from the data as an important factor affecting experimentation behaviour, and Garvin et al. also emphasises how through own example leaders can encourage employees to offer new ideas and options. 

These three building blocks overlap to some degree and reinforce one another. For instance, leadership behaviour helps in creating supportive learning environment and this supports managers and employees in creating and defining concrete learning processes and practices. Furthermore, concrete processes support leaders behaviour in a way that fosters learning and through own example cultivates that behaviour to others. 

Supportive leadership behaviour alone is not sufficient for guarantee organisational learning. Garvin et al. (2008) emphasise how organisations are not monolithic and managers should be sensitive to differences in culture, department and units. In addition to cultural differences, learning requires clear and targeted processes and practices. Furthermore, learning should be considered as multidimensional, thus organisational forces should not be solely focused on a single area but to consider presented building blocks as a whole. 

\subsection{Leadership behaviour supporting experimentation behaviour and learning}

Conventional leadership behaviour focuses on internal activities within the team, whereas innovative team leader needs various set of skills and approaches in order to encourage developing and growing of teams and individuals. For instance, according to Barckzak and Wilemon (1989) leaders of innovative teams utilise wide range of familiar and unfamiliar techniques in order to accomplish the team objectives, whereas leaders of operating teams use only a few familiar techniques. Even though in this study innovation teams were not studied, similar elements of developing by experimenting and encouragement for that may be recognised, when dealing with new tasks and developing something which result is uncertain. 

Also Bujis (2007) states how leaders dealing with uncertain and new innovations should stay certain about uncertainties and provide a safe environment and encourage employees to work on current task comfortably. Thus, high level of tolerance for dealing with different states of minds and various personal feelings is required from a leader. 

Leaders and their behaviour have great influence on the creativity and innovation ability of employees (Mumford et al. 2002).

In order to encourage creativity and experimenting in teams, leaders should lead by example and act as role models. Leaders should consider their own behaviour and actions in a way that stimulates employees to new and innovative, creative approaches to problems. In addition, they can even request creative and innovative solutions form the team, which may lead to better results in creativity of individuals (Amabile et al. 2002). Lack of time for creativity and developing is not usually considered as part of work, thus time allocated for that is needed, as was also found in the empirical part of the thesis.   (Waldman, mum for, amiable, khaire)

Big ideas do not hatch overnight and creative thinking requires time. Leaders should allow team members time to think creatively, as according to studies under pressure creativity actually falls into decline even though individuals may feel more creative, yet actually they are perhaps working more and getting things done. According to this study, employees were clearly less creative while time pressure increased. (Amabile et al 2002) 

Furthremore, leaders can assist their employees by recognising times with high pressure, and allowing employees to focus on certain thing at a time, leaving the expectations of creativity and new ideas into the future moment, when time pressure has decreased. On the other hand, if creativity is required under stress, leader should transparently explain the importance and reasons behind the strict schedule and required goals. Thus an employee may relate to the problem at hand and engage better at his work. Indeed, helping people to understand the importance of work is essential especially under high time pressure. (Amabile et al. 2002)

Through encouraging employees in risk-taking and making mistakes, leaders are likely to boost innovation. (Farson and Keyes, 2002)

Failure as a part of innovation and development process begins to be generally recognised and approved (Farson and Keyes, 2002). Succeeding companies even thrive for failure in order to learn fast and find the best practices and business models. For instance, credit company Capital One conducts continually large amount of market experiments. They now most of the tests will not pay off, yet they also know how much can be learned about customers and markets from failed tests in early phase of development.  (Farson and Keyes, 2002). Yet leaders fail in showing their employees the support and tools for failing fast and early enough. Failing in a personal matter remains a difficult subject, as failing never feels exceptionally great, and often employees still consider failed work as failing personally. (Farson and Keyes, 2002) 

Failure-tolerant leaders put effort on explaining to employees how important part failure is to the development process as a whole, and how failing actually refers to a point where surprising, failed outcomes are not reflected and further analysed in order to learn. Permorming accordingly, admitting own failures and not chasing anyone to blame, failure-tolerant leaders encourage failure, lower the threshold and ease the fear of failing of employees. (Farson and Keyes, 2002)

Naturally management need to take seriously issues about safety and health, yet most of the failures should be seen as opportunities for growth. Furthermore, failure-tolerant leaders treat success and failure similarly, analysing and reflecting the outcomes in order to grow the intellectual capital of the team, including experience, knowledge and creativity. Other characteristics of failure-tolerant leaders are being rather collaborative than controlling, listening carefully, seeing the bigger picture, asking questions and focusing on the development and future rather than blaming on mistakes. In addition, in order to gain empathy and trust among employees, leader should admit their own mistakes, as it shows self-confidence and honesty, assisting in forming closer ties with employees. Vulnerability and transparency play a major role in trustworthy relationship between leader and employees. (Farson and Keyes, 2002)

Through the green light given and their own example leaders can change the focus from success and failure into thinking in terms of learning and experience. (Farson and Keyes, 2002)


As Amabile and Khaire (2008) draw a poetical picture how leader cannot manage creativity, but manages for creativity. Furthermore, they suggest that culture that fosters creativity includes leadership that enables collaboration, enhances diversity, encourages ideation, maps the stages of creativity to different needs, accepts inability and utility of failure and motivates employees with intellectual challenges. 

In experimentation process, employees need to contribute imagination, and this may require new kind of encouragement for creativity from the leaders. (Amabile and Khaire, 2008) Much success rises from employee's own initiatives, which results from wide amount of autonomy at work. 

Change from authority-based leadership to collaboration with employees has occurred in literature and in practice (Amabile and Khaire, 2008; Farson and Keyes, 2002). 

A culture of creativity can be fostered in an organisation through opening the organisation to diverse perspectives and openness to various ideas. This calls for safe environment for employees to share their thinking from different fields of expertise. Furthermore, engouraging passion and knowledge of an employee is likely to result in more creative action at work,  (Amabile and Khaire, 2008)


Role models 
Lead by example
Allow experimenting 
Allow resources (material: Katz and Allen 1988)
Provide clear focus (Barckzak and Wilemon 1989)
Time (Gruber and Davis 1988)
encourage risk-taking (Sethi et al. 2001) 

Waldman and Bass 1991

Mumford et al (2002)

Hohn, 2000

Amabile et al. 2004

Khaire, 2008


 Tolerance for ambiguity 
 
 Ambiguity is often perceived by individuals when lacking sufficient cues to structure a situation, and usually arises from novelty, complexity or unsolvability of situation at hand (Budner 1962). 
 
\section{Practices and structures for developing and experimenting}

According to Amabile et al. (2002) clear time should be allocated for developing especially when the aim is to flourish idea generation, creativity, learning and experimentation of new concepts. Time pressure should be minimal in order bright ideas to glow as cognitive processing requires time of an individual and team. Yet, no sense of urgency leads employees easily to auto-pilot mode, in which routine tasks are performed without further thinking and analysing. Thus, creative time for playing with ideas, brainstorming, learning and experimenting should be allocated in an organisation in order truly new things to develop. Shared goals are once more essential in engaging team members to play with ideas and feel more motivated in developing their work.\subsection{Failing}

Failing is widely considered as essential part of learning (Farson and Keyes, 2002; )


Discussioniin

Relation between personal characteristics such as age, study background or education and experimentation behaviour was not studied in this thesis, yet some correlations were recognised, such as employees who had more working experience were more confident in performing experiments and they did not consider failing as a personal failure, but a reflective opportunity for learning. 
	\end{document}  